---
title: "Vertical Disintegration and the Shifting Boundary of the Farm Business: Implications for Agricultural Productivity and Market Structure"
execute:
  echo: false
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    page-layout: full
    df-print: paged
# a workaround for broken docx, see here https://github.com/quarto-dev/quarto-cli/issues/7321
  # docx:
  #   tbl-cap-location: bottom

bibliography: references.bib
---

```{r}
#| output: false
library(tidyverse)
library(modelsummary)

theme_set(theme_minimal())

source("projects/aaea-conf-2025-agser/agser.R", local = (agser <- new.env()))
glance.lmdf <- agser$glance.lmdf
tidy.lmdf <- agser$tidy.lmdf


paperdata <- readRDS(agser$opath$paper)
```

# 1. Introduction

Identifying the sources of productivity growth in the agricultural sector of the US economy is a core issue for researchers, policymakers, and producers.
Higher productivity improves competitiveness in international commodity markets, lowers the cost of food for American households, increases the welfare of farm families and agricultural workers, and strengthens the economic vitality of rural communities.
In this paper we study the role played by specialized service providers in shaping agricultural productivity growth.

One of the most robust stylized facts is the secular increase in farm total factor productivity (TFP).
Since 1948, farm output nearly tripled while use of farm inputs has remained flat [@usda-ers2025; @wang2024].
In practice, however, TFP is a residual—it is the increase in output that an aggregate production function specification is not directly measuring.
Allocating the observed TFP growth to specific sources is therefore vitally important for the promulgation of effective agricultural policy.

This growth has mainly been attributed to technological improvement, primarily fueled by R&D investments from public and private sectors, enhanced by public infrastructure, extension, and technology spillover from other sectors or neighboring regions [@alston2010; @huffman2006; @fuglie2007; @plastina2012; @wang2013].

These explanations may be incomplete, however, as they tend to ignore the evolving structure of agricultural production.
It is well-documented that farms have undergone significant consolidation over the past century.
The number of farms has fallen while the average size has increased with a heavy skew toward the very largest operations.
Less well studied has been the vertical disintegration of agricultural production characterized by the increasing share of production activities undertaken by specialist service providers rather than by farm operations [@dunn2017].

Despite the importance of the agricultural services sector to the farm economy, because of gaps in the federal statistical system, relatively little is known about how this industry has evolved over the past three decades.
@babkin2022 document establishment, employment, payroll, and wage dynamics for businesses that provide support services to crop production.
They demonstrate that agricultural support service is a uniquely dynamic and rapidly evolving sector of the American economy.
Although rates of entry have fallen substantially, large incumbent firms have tended to grow in size.
Furthermore, the real annual mean wage to nonagricultural workers employed in the sector has increased by nearly \$10,000 over the 1992-2016 period, indicating sizable growth in labor productivity.
Nonetheless, standard measures of economic dynamism, a principal source of productivity growth, have all fallen, which may have implications for innovation and the long-term health of this industry and agricultural production more generally.

The relationship between service use and agricultural productivity is a priori ambiguous.
At the national level, there is a strong evidence of substitution towards agricultural services, implying that it is an optimal choice by farm operators.
But farmers do not necessarily optimize for agricultural productivity, and rather for profitability of the farm business or even overall utility of the farm household.
Multiple factors can make switching to an external service provider preferable even if it leads to lower yields.
Among these are the time spent supervising hired laborers, the need to acquire and maintain agricultural machinery and risks of under-utilization over its lifespan, and the opportunity to allocate farm owner's time to activities outside of farm operation.
On the other hand, external service provider not be familiar with all the nuances of a specific farm operation, and additional inefficiencies are also introduced by the classic principal-agent friction.
In this paper, we use the administrative LBD data to investigate the relationship between availability agricultural service providers and farm productivity at a local market level.

Our findings for the agricultural sector are complementary to the rich literature studying the questions of vertical desintegration in the context of outsourcing in manufacturing and services.
It has been shown both theoretically and empirically that outsourcing is not always associated with increases in productivity [@heshmati2003; @görg2004; @lahiri2016].
Outsourcing is expected to involve production cost savings relative to internal production because outside suppliers benefit from economies of scale, smoother production schedules and centralization of expertise, but the choice between continued internal production or an outsourcing decision requires other considerations than pure production cost differences.

Theoretical literature suggests that the outsourcing decision can be viewed through multiple lenses.
Transaction cost economics argues that the optimal form of organization is primarily a function of the characteristics underlying a given exchange between a provider and its client, and that unfavorable exchange conditions can increase the cost of writing enforceable contracts [@coase1937; @williamson2010].
According to the resource-based view of the firm, outsourcing allows clients to gain access to specialized resources and capabilities residing in the provider firms and to focus more on core competencies [@barney1991].
From the perspective of the agency theory, client firms (principals) employ providers (agents) to perform certain functions or processes on their behalf, and the expected returns can be lowered by the risks of opportunism and moral hazard [@ross1973].
Multiple other theoretical approaches exist, but regardless of model predictions, realized effects of outsourcing on firm performance depend greatly on the implementation of the outsourcing strategy by firm management.

Empirically, a range of studies from different national contexts have shown that outsourcing can have positive [@bardhan2006], negative [@broedner2009] or no significant effects [@bengtsson2009] on firm performance.
An important observation is that empirical studies use a wide array of performance indicators, including financial metrics (return on assets, profit margins, sales), labor productivity and indirect measures such as customer satisfaction or rates of innovation.

In this study, we measure the scale of activity of the agricultural service providers in a county using total payroll of the LBD establishments in the NAICS 1151 industry.
These data offer multiple advantage over the alternative of using expenses reported in the Census of Agriculture (COA).
First, COA is a survey over a sample of farms with an overall response rate of 61% in the 2022 round [@usda-nass2024] and lower rates for individual questions.
LBD, on the other hand, is based on administrative IRS records and includes every active establishment with positive payroll, providing universal coverage and highest data accuracy.
Second, payroll expenditures paid by service providers to workers better track the quantity of labor input used in production than the amounts paid to providers by farmer.
Finally, the use of microdata resolves the issue of cell suppression and allows us to calculate custom indicators characterizing local markets, such as measures of market concentration.

Considerable number of studies have explored both theoretically and empirically the role of market power of food manufacturers and retailers [@sexton2013; @sheldon2017; @smith2015; @gaudin2018; @ma2019; @sexton2018; @moschini2025].
On the upstream side of the farm, the literature is much smaller and has been mostly focused on increasing market concentration of seeds and agricultural chemicals suppliers [@macdonald2017; @fuglie2012; @hendrickson2017; @wilson2010].
Our paper offers new insights into the structure of the agricultural services market and paves the way for a more nuanced micro-level analysis of the effect of imperfectly competitive pricing on productivity growth accounting.

The article is organized as follows.
In Section 2 we provide a set of stylized facts characterizing the trends in farm productivity, expenditures and the agricultural services sector.
Data and methodology are presented in Section 3.
Results of the descriptive and regression analyses are in Section 4.
Section 5 concludes with a summary of our findings and a discussion of future work.

# 2. Background

Estimates of the agricultural TFP show a steady increase of real output and productivity with a stable level of real input since 1948, and the pattern also holds over the recent two decades (@fig-tfp-1).
Within the inputs category, decreases in real labor expenditures have been compensated by an increase of spending on intermediate inputs, with capital spending being mostly flat, although recently the pattern has weakened (@fig-tfp-2).
Decomposition of intermediate input expenditures reveals that the growth was primarily driven by the increase of the real use of pesticides, the rest of the spending on intermediates staying mostly flat or declining in real terms since 2002 (@fig-tfp-3).
One of the main expenditure categories attributable to service providers (the other being contract labor) is the intermediate services, and it remained flat over the last two decades.
This lack of growth in the use of constant-efficiency units of intermediate services might be caused by underestimation of quality improvement as discussed in the introduction.

```{r}
#| label: fig-tfp
#| fig-cap: "Agricultural productivity decomposition, 2002 = 100%"
#| fig-subcap:
#|   - "Output, input and TFP"
#|   - "Capital, labor and intermediates"
#|   - "Intermediate inputs"
#| layout-ncol: 3

paperdata$tfp %>%
  filter(name %in% c("output", "input", "tfp")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

paperdata$tfp %>%
  filter(name %in% c("capital", "labor", "int")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

paperdata$tfp %>%
  filter(str_starts(name, "int_")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

```

Decomposition of labor expenditures into hired and contract is not available in the published agricultural productivity accounts, but it is informative to consider spending reported in the Census of Agriculture, deflated to 2023 dollars using the GDP price index (@fig-exp-labor).
Contract labor expenditures grew by %88 from 2002 to 2022 in real terms, double of the %44 growth rate of farm spending on hired labor, strongly indicating a shift towards the use of specialized labor service providers.
It should be noted that these trends are not directly comparable to an approximately 10% drop in the quality-adjusted labor use shown in [@fig-tfp-2], primarily because the latter series is deflated using a quality-adjusted wage index which outpaced the general price growth rate reflected in the GDP price index.

```{r}
#| label: fig-exp-labor
#| fig-cap: "Labor expenditures"
#| fig-subcap:
#|   - "In 2023 dollars"
#|   - "Normalized, 2002 = 100%"
#| layout-ncol: 2

d <- paperdata$exp_labor

d %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

d %>%
  ggplot() +
  geom_line(aes(year, value_norm, color = name)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0.8, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))
  

```

Another indication of continuing substitution of agricultural activity from farms to specialized providers can be drawn from the trends in the value of agricultural machinery owned by them ([@fig-asset]).
These series were built from the detailed industry-by-asset data from Fixed Asset Accounts program of the Bureau of Economic Analysis, deflated to 2023 dollars using the GDP price index.
Although the agricultural services industry is combined with forestry and fishing, the share in the reported stock of tractors and other agricultural machinery held by them is likely low.

```{r}
#| label: fig-asset
#| fig-cap: "Value of agricultural machinery"
#| fig-subcap:
#|   - "Tractors"
#|   - "Other agricultural machinery"
#|   - "Normalized, 2002 = 100%"
#| layout-ncol: 3

d <- paperdata$asset
pal <- scales::pal_hue()(2)

# numbers in text
# d %>%
#   filter(year %in% c(2002, 2022), asset == "other ag machine") %>%
#   mutate(pct = 100 * value / sum(value), .by = year) %>%
#   arrange(year, industry)

d %>%
  filter(asset == "tractors") %>%
  ggplot() +
  geom_line(aes(year, value, linetype = industry), color = pal[1]) +
  scale_x_continuous(name = "", breaks = seq(2002, 2022, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

d %>%
  filter(asset == "other ag machine") %>%
  ggplot() +
  geom_line(aes(year, value, linetype = industry), color = pal[2]) +
  scale_x_continuous(name = "", breaks = seq(2002, 2022, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

d %>%
  ggplot() +
  geom_line(aes(year, value_norm, color = asset, linetype = industry)) +
  scale_x_continuous(name = "", breaks = seq(2002, 2022, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

```

The stock of tractors owned by farms is much larger and growing in real terms, and that of service providers smaller and shrinking over time.
The stock of other agricultural machinery, on the contrary, is showing a very different picture.
While service providers hold a smaller stock in dollar terms, over the 20 years it has grown by 624% compared by only 62% growth of stock held by farms.
If in 2002 service providers held only 4.4% of the combined stock of other agricultural machinery, by 2022 this share has grown to 17.2%.
The difference in this trend compared to that of tractors - a more general purpose machinery - is consistent with the narrative of service providers focusing on more specialized processes required a more specialized type of asset.

Employment and payroll dynamics of the agricultural services provider sector has been documented in [@babkin2022], with relevant findings replicated in [@fig-agser-dynamics].
While the number of active establishments was under 10,000 and slowly decreasing over time with exit rates exceeding entry, the sector demonstrated steady growth of scale as measured by the total real payroll.
Establishment level dataset (Longitudinal Business Database) used to construct these statistics is assembled from administrative IRS records, which report payroll to agricultural and non-agricultural workers separately.
Consistent with the growth of farm expenditures on contract labor shown in [@fig-exp-labor], total payroll of agricultural workers employed by service providers has approximately doubled over the observed period.
Payroll of non-agricultural workers remained mostly flat, more in line with the pattern of farm spending on other agricultural services.

```{r}
#| label: fig-agser-dynamics
#| fig-cap: "Dynamics of the support activities for crop production sector (NAICS 1151)"
#| fig-subcap:
#|   - "Number of establishments"
#|   - "Establishment entry and exit rates"
#|   - "Total payroll, 2023 dollars"
#| layout-ncol: 3

# paperdata$agser_size %>%
#   filter(name == "pay_ag", year == 2012)
# paperdata$exp_labor %>%
#   filter(name == "exp_labcont", year == 2012)

paperdata$agser_size %>%
  filter(name == "est") %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_number(scale_cut = scales::cut_short_scale()))

paperdata$agser_size %>%
  filter(name %in% c("entry", "exit")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 0), legend.position.inside = c(0, 0))

paperdata$agser_size %>%
  filter(name %in% c("pay_nonag", "pay_ag")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

```

While growth rates of agricultural workers payroll and contract labor expenditures are very similar, the payroll is smaller in absolute magnitude (\$4 billion compared to \$8.5 billion in 2010).
It is not possible to explain the mismatch with certainty, but there are several important reasons.
First, the payroll reported here only covers the sector of support activities for crop production (NAICS 1151) and does not include support activities for animal production (NAICS 1152), which also serves to a smaller extent as a source of contract labor to farms.
Second, the payroll figure does not include payments to agricultural workers employed by business with primary NAICS activity code other than 115.
Finally, the numbers reported as the farm contract labor expenditures are the payments made by the farmers to service provider businesses that include business owner's profits, taxes, returns to capital and other non-labor expenses.

# 3. Methodology

*This section will be further expanded.*

## Data sources

-   Census of Agriculture, 2002-2022.
    Definition of county samples.
    Imputation of missing values.
    Approximation of land HHI.

-   Quarterly Census of Employment and Wages (QCEW).
    Publicly available NAICS 1151 data, subject to significant suppression at a county level.

-   Longitudinal Business Database, 2022-2017.
    Establishment level microdata constructed from administrative IRS tax records.
    Aggregated to county level observations.

## Regression models

Productivity, payroll, land and average land are always log-transformed.
Productivity is either sales per acre or yield (bushels per acre).

Static productivity without interaction:

$$
Productivity = Payroll + Land + AvgLand + LandHHI 
$$ {#eq-reg-static}

Static productivity with interaction:

$$
Productivity = Payroll + Land + Payroll * Land + AvgLand + Payroll * AvgLand + LandHHI 
$$ {#eq-reg-static-int}

Dynamic productivity, $\Delta Productivity(t) = Productivity(t+5) - Productivity(t)$ in logs:

$$
\Delta Productivity = Productivity + Payroll + Land + AvgLand + LandHHI 
$$ {#eq-reg-dynamic}

# 4. Results

A key contribution of our study is the use of the Longitudinal Business Database (LBD) to measure the activity of the agricultural services providers in NAICS industry 1151.
This dataset is only accessible through an approved research project in the FSRDC environment, but the best publicly available data source is vastly inadequate to perform the analysis of the relationship between farm and services activity.
@tbl-qcew-sup shows aggregation of county level QCEW data on NAICS 1151 establishments, employment and payroll merged with Census of Agriculture measures of harvested land, sales and the number of farms engaged in crop production in 2012.
Counties included in the "suppressed" column only have the number of service provider establishments reported, while employment and payroll are suppressed to protect business data confidentiality.
The 76% of counties with suppression host 38% of all agricultural services establishments.
And while these establishments are mostly small and constitute only 15% of total employment and 18% of total payroll, they operate in counties where more than half of all crop production is taking place.
<!--# TODO: Add a column with farm number for counties with no agser establishments, update the totals and pct_supressed -->

```{r}
#| label: tbl-qcew-sup
#| tbl-cap: Data suppression in 2012 QCEW.

paperdata$qcew_sup %>%
  datasummary_df(fmt = modelsummary:::fmt_identity())
```

Using the LBD, we can characterize the structure of the local agricultural services market.
@tbl-agser-distr reports the evolution of the number of establishments, payroll and market concentration in the NAICS 1151 sector at the county level from 2002 to 2017.
In line with the national aggregate trends reported by @babkin2022, the average number of establishments has been slowly declining while the average real payroll has nearly doubled from \$2.4 to \$4.0 million.
Notably, there is no evidence of an increasing local market concentration, as the mean and median payroll shares of the largest establishment in a county remained fairly stable around 73% and 78% respectively.
This finding is in sharp contrast to the trends of increasing market concentration in the agricultural production on the farm and in the adjacent markets along the production chain beyond the farm gate.
<!--# TODO: elaborate more on this, add farm distribution numbers and quote findings from the market structure literature -->

```{r}
#| label: tbl-agser-distr
#| tbl-cap: "County distribution of service providers"

paperdata$agser_distr %>%
  select(variable, year, nobs, mean, sd, `25%`, `50%`, `75%`) %>%
  arrange(variable, year) %>%
  datasummary_df(fmt = modelsummary:::fmt_identity())

```

Regression results for the estimation of the model in [@eq-reg-static-int] for all-crop sample in 2017 are presented in [@tbl-reg-static-crop].
Coefficients of lesser interest have been suppressed for confidentiality disclosure avoidance, showing only the sign and statistical significance.
This model shows there is a positive and statistically significant relationship between farm crop productivity and total payroll of the service providers operating in a county, and this effect persist controlling for land distribution variables.
A 1% increase in the payroll size of the services sector is associated with a 0.2% increase in crop sales per acre of harvested land.
We interpret this as an evidence of the positive impact of availability of service providers on farm productivity.
The effect is stronger in counties with more total acreage and smaller average farm size, consistent with a story of service providers becoming more effective when serving many small farms at scale.

```{r}
#| label: tbl-reg-static-crop
#| tbl-cap: "Regression of crop sales per acre on service provider payroll, 2017."

models <- paperdata$reg_static %>%
  filter(commodity == "crop", year == 2017)

fits <- unique(models$model) %>%
  map(\(l) filter(models, model == l))

modelsummary(fits, output = "huxtable")
```

The advantage of using dollar crop sales per acre as the measure of agricultural land productivity is the ability to aggregate all crop production into a single number.
But it comes at a cost of obscuring the variation of productivity and its responsiveness to availability of service providers across different crop categories.
In particular, vegetables, fruits and horticulture tend to return higher sales per acre than grains while also being more labor intensive to grow.
Since our measure of service provider size - payroll to employees - most directly reflects the labor component of support activity, counties with sales dominated by vegetables, fruits and horticulture are more likely to have service providers with high payrolls regardless of farm productivity, leading to an upward bias in the estimated coefficient.
As we do not observe in the data which crops service providers are involved with, our approach to address this issue is to look at yields of three major homogeneous grain crops - corn, soybeans and wheat - measured in bushels per acre and to restrict the samples to counties with positive production of respective crops.
In a symmetric fashion, this approach would attenuate the coefficient of interest because the payroll is potentially affecting productivity of commodities other than the one which yield is being used as the outcome in the model.
Measuring output in bushels instead of dollars also avoids the issue of unobserved unit price variation.

[@tbl-reg-static-com] presents payroll coefficients for the static productivity model without interaction estimated over corn, soybeans and wheat samples in 2002, 2007, 2012 and 2017.
A 1% increase in agricultural service payroll is associated with a 0.02-0.06, 0.01-0.03 and 0.02-0.05 percent increase in corn, soybeans and wheat yields, respectively.
While the magnitude of these estimates is smaller than in the sales-per-acre model, they are always statistically significant and likely underestimated as discussed above.

```{r}
#| eval: true
#| label: tbl-reg-static-com
#| tbl-cap: "Payroll coefficient estimates"

models <- paperdata$reg_static %>%
  filter(model == "logyield ~ logpay + logland + logavgland + hhi_land",
         term == "logpay",
         commodity %in% c("corn", "soy", "wheat"))

models %>%
  mutate(value = paste0(round(estimate, 3), significance, " (", round(std.error, 3), ")")) %>%
  pivot_wider(id_cols = year, names_from = commodity) %>%
  datasummary_df(fmt = 0)


```

The effect of the service payroll on yields varies by the distribution of harvested land, as identified by the interaction terms in [@eq-reg-static-int].
Full set of estimates are reported in the Appendix Table ...
<!--# TODO: add the table --->, and for easier interpretation, [@fig-me] shows marginal effects of payroll evaluated at high and low quartiles of the respective land variables.
Large magnitudes of variation of the marginal effect across time and commodities suggest that heterogeneity might be present, but statistical significance of the estimates is generally low, and no strong patterns emerge.

```{r}
#| label: fig-me
#| fig-cap: "Marginal effect of payroll on yield"
#| fig-subcap:
#|   - "At different levels of land"
#|   - "At different levels of average land"
#| layout-nrow: 2
#| fig-asp: 0.4

paperdata$reg_me %>%
  filter(commodity != "crop", atvar == "logland") %>%
  rename(land = atval) %>%
  ggplot() +
  geom_line(aes(year, value, color = land)) +
  facet_wrap("commodity", scale = "free_y", nrow = 1) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  scale_y_continuous(name = "") +
  scale_x_continuous(name = "", breaks = seq(2002, 2017, 5)) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))


paperdata$reg_me %>%
  filter(commodity != "crop", atvar == "logavgland") %>%
  rename(avgland = atval) %>%
  ggplot() +
  geom_line(aes(year, value, color = avgland)) +
  facet_wrap("commodity", scale = "free_y", nrow = 1) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  scale_y_continuous(name = "") +
  scale_x_continuous(name = "", breaks = seq(2002, 2017, 5)) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

```

We recognize that the cross-sectional estimates presented here are possibly subject to endogeneity bias.
While productivity enhancing effect of agricultural service provider availability is a reasonable explanation for the observed positive relationship, the effect might go in the opposite or, more likely, in both directions: The most productive farms are more dynamic and innovative, and create demand for specialized service providers to appear in their local markets.
In this study, we begin to untangle this relationship by analysing the dynamic response of farm yields to provider availability.
Model presented in [@eq-reg-dynamic] evaluates whether counties with more agricultural provider activity demonstrated higher growth rate in yields over the subsequent five year period.
<!--# TODO: explain why we control for base year productivity in these regressions --> Payroll coefficient estimates are collected in [@tbl-reg-dynamic].
All statistically significant coefficients in these models are positive, supporting the hypothesis that agricultural service providers increase farm productivity.
To illustrate with the largest coefficient, crop-growing counties that had a 1% larger service sector experienced a 0.05% growth rate in corn yields.

```{r}
#| eval: true
#| label: tbl-reg-dynamic
#| tbl-cap: Payroll coefficient estimates

models <- paperdata$reg_dynamic %>%
  filter(model == "diff_logyield ~ logpay + logyield + logland + logavgland + hhi_land",
         term == "logpay")

models %>%
  mutate(value = paste0(round(estimate, 3), replace_na(significance, ""), " (", round(std.error, 3), ")")) %>%
  pivot_wider(id_cols = year, names_from = commodity) %>%
  datasummary_df(fmt = 0)


```

# 5. Discussion

In this paper, we present new estimates of the relationship between farm productivity and agricultural support services using both cross-sectional and time-series variation.

Results using publicly available data from the Quarterly Census of Employment and Wages (QCEW) indicate that in the cross section, greater access to agricultural service providers is associated with higher farm productivity.
This relationship appears to be highly heterogeneous, varying with the agricultural commodity and the service provided.
It also depends on the size of farm operations.
However, a noteworthy limitation of using publicly available information on service providers is the high levels of data suppression that occur to maintain disclosure avoidance standards.
Particularly when using county-level data at the six-digit NAICS industry level, the plurality of observations is suppressed.
Further, the treatment of agricultural and non-agricultural labor varies across statistical agency, data program, and administrative purpose.
Therefore, a significant innovation of this study is the use of the unsuppressed Federal administrative data on establishment payroll tax filings from the Internal Revenue Service contained in the Longitudinal Business Database for businesses in NAICS industry 1151 (Support Activities for Crop Production).

Using the LBD and the Census of Agriculture for 2002, 2007, 2012 and 2017, we find that 1% larger agricultural support services sector in a county, as measured by total payroll, is associated with 0.2% higher farm sales per acre, controlling for farm characteristics.
The relationship is also strong for corn and wheat yields, but not for soybean yields.
Positive effect on productivity is amplified in counties with larger farmland acreage.
Leveraging the longitudinal nature of the LBD, we find that counties with higher service provider availability experience higher productivity growth over the subsequent 5-year period.

## Implications for productivity accounting

TODO: revise to fit the concluding section

Decomposition of farm input usage reveals that large reductions in use of land and labor have been offset by increases in capital and intermediate inputs, the latter showing an increase of 109 percent from 1948 to 2021 [@wang2024].
Just as economies of scale may accrue to larger farms, support service providers may be able to increase capital intensity and adopt new technologies more quickly.
By focusing on a narrowing set of production activities, service providers may also experience significant gains from specialization.
It might well be that existing quality adjustments to the cost of intermediate inputs are underestimated.
Specifically, returns to scale and specialization may make agricultural service providers more productive than farmers at on-farm production activities, increasing outputs while decreasing costs.

A key component of modern agricultural productivity accounting is quality adjustment.
TFP growth is calculated as the difference between the growth rates of output and inputs, and aggregation across different commodity and expenditure types necessitates the use of dollar denominated nominal quantities, which are in turn deflated to real terms using a combination of price indices.
The price index is constructed to account both for relative price variation over time and geography for factors where quality does not change (e.g. energy) or quality measurement is not available (farm products), and for changes in quality coming from either intrinsic quality improvements (a pesticide becoming more potent) or substitution between different types (different education attainment of workers).
The classic argument of @jorgenson1967 is that the estimate of TFP or “the measure of our ignorance” can be explained away by accurately accounting for changes in all the inputs, and by doing so we can better understand the sources rather than just the magnitude of growth.
Indeed, input quality changes in labor, capital, and intermediate inputs explain 0.11, 0.04, and 0.04 percentage points, respectively, of the 1.46 percentage points of annual agricultural output growth [@wang2024, p. 26].

Depending on data availability, @wang2024 use two methods to calculate quality-adjusted price indices – hedonic and index number approach.
The index number method requires detailed and disaggregated data on prices, quantities and composition of characteristics of a particular input.
The hedonic method applies regression analysis to estimate year dummy coefficients and can be used with less available data.
While both methods allow for a more accurate measurement of input volumes in constant efficiency units, it is important to recognize their limitations that may lead to biases.

Two primary farm expenditure categories associated with agricultural service providers are contract labor and purchased machine services.
Nominal contract labor expenditures are deflated with a hedonic price index of agricultural worker earnings, estimated from the National Agricultural Workers Survey (NAWS) that includes an indicator variable of whether the worker was employed by a contractor [@wang2025].
This approach relies on the assumptions of representative sampling and correct model specification, with potential biases for price index estimation in both directions.
Purchased machine services are considered a substitute for own-capital input, and the same price index is applied for quality adjustment of both expenditure types, possibly underestimating the quality of the work performed by a specialized provider for the same nominal dollar cost.

While the national productivity accounting framework can be used to study the effect of agricultural service providers on the aggregate farm productivity, the accuracy of this approach would only be as good as the accuracy of the quality adjustment procedures, which currently suffers from the aforementioned limitations.
In this paper, we contribute to the agricultural productivity discussion by generating estimates of the relationship between farm productivity and scale of provider operation identified from the spatial variation across United States counties.
This approach allows us to draw further insights on the heterogeneity of this relationship over different crop types and farm land distributions.

The theoretical foundation of productivity accounting also relies on the assumption of perfectly competitive market equilibria where output prices are equal to marginal costs of production, and input factors are paid their respective marginal product (see [@hulten2010, p. 1023] for a brief overview).
Under this assumption, cost shares are equal to the corresponding output elasticities and can thus be used in the construction of the Divisia-Tornkvist growth indices.
In practice, however, product and input markets are often subject to varying degrees of market power on either side of the trade, introducing distortions to the calculation of quality adjusted quantities.
For illustration, if agricultural service providers are increasing their trading power in the local markets over time, service price will be increasing over its marginal product, but for an econometrician estimating a hedonic price index this increase will be indistinguishable from an increase in the unobserved quality of service, leading to biases in the estimates of constant efficiency quantity and TFP.

# References

::: {#refs}
:::
