---
title: "Vertical Disintegration and the Shifting Boundary of the Farm Business: Implications for Agricultural Productivity"
author:
  - Anton Babkin
  - Richard A. Dunn
date: today
execute:
  echo: false
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    page-layout: full
    df-print: paged
  docx:
    link-citations: true
  # a workaround for broken docx, see here https://github.com/quarto-dev/quarto-cli/issues/7321
  #   tbl-cap-location: bottom

bibliography: references.bib
nocite: "@*" # include all references from the .bib file
---

```{r}
#| output: false
library(tidyverse)
library(modelsummary)
library(tinytable)

theme_set(theme_minimal())

source("projects/aaea-conf-2025-agser/agser.R", local = (agser <- new.env()))
glance.lmdf <- agser$glance.lmdf
tidy.lmdf <- agser$tidy.lmdf


paperdata <- readRDS(agser$opath$paper)
```

# 1. Introduction

Agricultural production in the United States has undergone significant vertical disintegration over the past half century [@dunn2017].
On-farm production activities – soil preparation, planting, cultivation, harvesting, and pre-market processing – are increasingly undertaken by specialist providers who supply labor, capital, and management services.
The choice to outsource production activities, while profit or utility maximizing for the farm business, may not be productivity enhancing.
Evidence from other sectors of the economy indeed suggest that the relationship between the outsourcing of production activities and productivity is both theoretically and empirically ambiguous. 

Despite the importance of agricultural productivity to the national, regional, and local economies of the United States and its central role in shaping domestic energy, environmental conservation, rural economic development, international trade, and immigration policy, the effect of agricultural service providers on agricultural productivity has not been adequately documented.
Therefore, in this article we estimate the relationship between access to agricultural service providers and crop yields using Federal administrative data that overcomes the suppression and survey non-response issues that exist in other data sources.

The agricultural services sector has been expanding at a faster rate than the farm sector.
Between 1978 and 2022, the real value of sales by agricultural service providers increased 263 percent, compared to just 12 percent for farms (Dunn, 2024) <!--# TODO: Dunn, Richard A. 2024. "2022 Economic Census: The growing contribution of support services to U.S. agricultural production". USDA Economic 
Research Service, Charts of Note. -->.
The increased reliance on agricultural service providers has affected patterns of farm labor expenditures and ownership of agricultural machinery.
According to the Census of Agriculture, the share of contract workers in farm paid labor expenditures went from 11.6% in 1982 to 19.5% in 2022.
The share of tractors and other agricultural machinery owned by service providers over the same period grew from 6.1% to 14.1%, as reported by the Bureau of Economic Analysis (BEA) in its Fixed Asset Accounts data.
Thus, by 2022 a substantial proportion of farm labor and machinery was managed by non-farm businesses, which is especially remarkable considering that the number of agricultural service establishments is less than 1% of the number of farm businesses.

Despite these long-term trends, the potential implications of outsourcing agricultural production activities to non-farm businesses is widely understudied.
Of particular note, we are unaware of any published articles that estimate the effect of outsourcing on agricultural productivity in the United States.
Further, the existing literature that considers other sectors or countries is contradictory.

Theoretical work indicates that the decision by profit-maximizing firms to outsource production activities can either increase or decrease productivity [@heshmati2003; @görg2004; @lahiri2016].
Gains to specialization and the economies of scale available to service providers would tend to improve productivity by increasing output and/or lowering costs.
On the other hand, the absence of firm-specific knowledge and inefficiencies arising from moral hazard in the principal-agent relationship between firms buying and selling production services would tend to lower productivity.
These trade-offs can yield cases where a profit-maximizing firm opts to outsource, accepting lower productivity because of sufficient cost-savings.

Agricultural operations introduce an added complexity that is generally ignored in existing theoretical models of outsourcing, namely that the overwhelming majority of farms in the United States are family businesses.
Households are utility-, not profit-maximizers, and the farm business is often only one of several income sources for farm families.
The decision to outsource production activities may allow for a greater allocation of time to off-farm labor opportunities or simply more leisure, increasing the incentive to accept lower on-farm productivity.
Alternatively, allocating some production activities, including management and oversight of labor or capital, to a third party may allow farm operators to more effectively plan and manage other aspects of the farm business, ultimately increasing overall farm productivity.
The countervailing effects of time entering both the utility function and income constraint of farm households only increases the potential ambiguity between outsourcing and productivity in agricultural production.

The theoretical ambiguity between outsourcing and productivity is reflected in the conflicting set of empirical results that examine manufacturing and service industries, with findings that alternatively show that outsourcing can have positive [@bardhan2006], negative [@broedner2009] or no significant effects [@bengtsson2009] on firm performance.
<!--# TODO: add one more of each -->The empirical literature that considers the relationship between productivity and outsourcing specifically in agriculture is altogether lacking with one notable exception.
@lu2022 examine the effect of agricultural mechanization services in China.
Using a 2018 survey of rural households, they find positive effects in plowing, transplanting, and harvesting services, and a negative effect in pesticide spraying.
To the best of our knowledge, however, no studies exist on this topic for agriculture in the United States.

Studies of the agricultural services sector also offer mixed findings on the potential relationship between outsourcing and agricultural productivity in the United States.
Using the National Agricultural Workers Survey (NAWS), @wang2025 find that farm workers employed by contract labor providers receive lower wages than workers hired directly by farms.
This would suggest that labor productivity is lower for contract labor.
Yet, @babkin2022 find that the real annual mean wage to nonagricultural workers employed by establishments providing support services to crop production increased nearly \$10,000 between 1992 and 2016, indicating potentially significant growth in the productivity of the workers managing and directing farm workers.

Given both the theoretical ambiguity and the lack of consistent empirical findings within a relatively thin literature, we investigate whether greater access to agricultural service providers increase field crop yields on U.S. farms.
We find that in every census year between 2002 and 2017, counties with larger total payroll in the services sector have higher crop yields, with elasticities of 0.02-0.062 for corn, 0.007-0.027 for soybeans and 0.024-0.048 for wheat.
This relationship varies substantially by the total acreage of harvested cropland, getting consistently stronger in larger counties for wheat and weaker for corn, suggesting different scale economies.
We do not find strong evidence in support of a hypothesis that service providers have bigger effect on productivity in counties with larger number of smaller farms, if anything, the opposite is more likely.

We also find that counties with more services experienced greater growth yields over subsequent 5-year periods, allowing for a more causal interpretation of the relationship between service provider use and agricultural productivity, but the effect tends to get weaker in later years.
This trends coincides with the falling business dynamism in the agricultural services industry reported by @babkin2022.
One possible interpretation is that the productivity gains from outsourcing are reaching their limits as the industry matures, and the associated costs begin to outweigh the benefits.

These results have a number of important implications for productivity growth accounting.
First, usage of identical price indices used for owned machinery and custom work expenditures likely underestimates the quality-adjusted growth rate of the latter and overestimates farm TFP growth.
Second, such adjustment will be more accurate if crop and acreage heterogeneity of productivity effect is taken into account.
Third, if service providers exert market power over workers more than farms do, expenditure shares overestimate output elasticity of labor, leading to overestimation of farm TFP .
We expand upon each of these in the discussion section.

The article is organized as follows.
In Section 2 we provide a set of stylized facts characterizing the trends in farm productivity, expenditures and the agricultural services sector.
Data and methodology are presented in Section 3.
Results of the descriptive and regression analyses are in Section 4.
Section 5 concludes with a summary of our findings and a discussion of future work.

# 2. Background {#sec-background}

Farm owners can contract outside firms to perform almost any part of the agricultural production process, from narrow and specific tasks such as seed bed preparation or crop packing to overall management of farm day-to-day operations.
The focus of this study is the industry defined by the North American Industry Classification System (NAICS) as Support activities for crop production (code 1151) which we would call agricultural services for simplicity.
At a finer level of classification this sector includes Cotton ginning (115111), Soil preparation, planting, and cultivating (115112), Crop harvesting, primarily by machine (115113), Postharvest crop activities (except cotton ginning) (115114), Farm labor contractors and crew leaders (115115), and Farm management services (115116).

The agricultural services industry is relatively small - about 10,000 businesses nationwide - and all but the most basic analysis of it is problematic with public data.
@dunn2017 and @babkin2022 provide a detailed coverage of the statistical challenges for the study of this sector.
The three publicly available datasets widely used to study specific industries are the Quarterly Census of Employment and Wages (QCEW) from the Bureau of Labor Statistics and the Census Bureau's County Business Patterns (CBP) and Business Dynamics Statistics (BDS).
As we describe later in this paper, a significant portion of the QCEW data is missing at a county level, although it can still be used to examine nationwide trends.
And both CBP and BDS do not cover about a half of the NAICS 1151 sector because they do not include establishments employing only agricultural workers.

@fig-agser-nat shows how the sector has evolved over the last thirty years using alternative data sources - QCEW and different revisions of the LBD.
The panels show the total number of employer establishments reporting 1151 as their primary industry and the real payroll paid to employees.
Although there are some disagreements, stemming from the different administrative sources of data (federal IRS payroll tax records for LBD and state unemployment insurance records for QCEW), the industry is characterized by two trends.
First, the number of establishments remained relatively stable or even started growing in recent years, contrary to trends in many other industries including farming.
Second, the size of the sector measured by the total real payroll has been steadily growing, roughly doubling over the observed period, and the growth was primarily driven by the increasing payroll to agricultural workers.

```{r}
#| label: fig-agser-nat
#| fig-cap: "Agricultural services industry (NAICS 1151)"
#| fig-subcap:
#|   - "Number of establishments"
#|   - "Total payroll in 2023 dollars"
#| layout-ncol: 3

paperdata$agser_nat %>%
  rename(`data source` = "src") %>%
  filter(name == "est") %>%
  ggplot() +
  geom_line(aes(year, value, linetype = `data source`)) +
  scale_y_continuous(limits = c(0, NA), name = "") +
  theme(legend.position = "inside", legend.justification = c(1, 0), legend.position.inside = c(1, 0))

paperdata$agser_nat %>%
  rename(`data source` = "src") %>%
  filter(name == "pay") %>%
  ggplot() +
  geom_line(aes(year, value, linetype = `data source`)) +
  scale_y_continuous(limits = c(0, NA), name = "", labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(1, 0), legend.position.inside = c(1, 0))

paperdata$agser_nat %>%
  filter(src == "LBD", name %in% c("pay_ag", "pay_nag")) %>%
  ggplot() +
  geom_line(aes(year, value, linetype = name)) +
  scale_y_continuous(limits = c(0, NA), name = "", labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(1, 0), legend.position.inside = c(1, 0))
```

Looking at the scale of service usage from the farm perspective, we see similar trends.
Drawing data from the Census of Agriculture, @fig-exp-agser shows two farm expenditure categories associated with crop services - custom work and contract labor - along with QCEW payroll of service businesses from 2002 to 2022.
Absolute levels of farm expenditures and service payroll are in comparable range, but can not be directly matched for several reasons.
First, farm expenditures are the payments made by the farmers to service provider businesses that include business owner's profits, taxes, returns to capital and other non-labor expenses.
Second, the payroll reported here only covers the sector of support activities for crop production (NAICS 1151) and does not include support activities for animal production (NAICS 1152), which also serves to a smaller extent as a source of contract labor to farms.
Finally, the payroll figure does not include payments to agricultural workers employed by business with primary NAICS activity code other than 115.
Nevertheless, the growth trends between the two data sources are remarkably close, consistently showing that the services sector has nearly doubled in size over two decades.

```{r}
#| label: fig-exp-agser
#| fig-cap: Farm expenditures on contract labor and custom work, 2023 dollars.
paperdata$exp_agser %>%
  ggplot() +
  geom_line(aes(year, value, linetype = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(1, 0), legend.position.inside = c(1, 0))

# growth rates
# paperdata$exp_agser %>%
#   filter(year %in% c(2002, 2022)) %>%
#   mutate(value_norm = value / first(value), .by = name)
```

To put these trends in the context of the farm sector, we compare service expenditures with analogous costs of labor and machinery employed by the farmer directly.
Leaving out the labor of farm family and other unpaid workers, paid worker expenditures between hired and contract labor can be cleanly compared within the Census of Agriculture.
@fig-exp-labor tracks farm paid labor expenditures from 1982 to 2022 in real dollar terms and as a normalized growth rate.
While farms spend much more on hiring workers directly than through service providers (\$43.3B compared to \$10.5B in 2022), the latter category was consistently growing faster than the latter over four decades, cumulatively increasing almost twice as much.

```{r}
#| label: fig-exp-labor
#| fig-cap: "Farm paid labor expenditures"
#| fig-subcap:
#|   - "In 2023 dollars"
#|   - "Normalized, 1982 = 100%"
#| layout-ncol: 2

# paperdata$exp_labor %>%
#   filter(year == 2022)

paperdata$exp_labor %>%
  ggplot() +
  geom_line(aes(year, value, linetype = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

paperdata$exp_labor %>%
  ggplot() +
  geom_line(aes(year, value_norm, linetype = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))
```

Trends in the use of capital can not be as easily drawn from the Census.
While spending on outsourced machinery can be taken from custom work expenditures, the flow value of owned capital services is not observed and is typically inferred from the value of capital stock in relevant applications.
Instead, we can compare the stock of agricultural capital owned by farms and service providers using detailed industry-by-asset data from Fixed Asset Accounts program of the Bureau of Economic Analysis ([@fig-asset]).
Although the agricultural services industry is combined with forestry and fishing, the share in the reported stock of tractors and other agricultural machinery held by them is likely low.

```{r}
#| label: fig-asset
#| fig-cap: "Value of agricultural machinery"
#| fig-subcap:
#|   - "Real value of tractors and other agricultural machinery"
#|   - "Normalized, 1982 = 100%"
#| layout-ncol: 2


# numbers in text
# paperdata$asset %>%
#   filter(year %in% c(1982, 2002, 2022)) %>%
#   mutate(value = value / 1e9) %>%
#   pivot_wider(id_cols = c(year, asset), names_from = industry)

paperdata$asset %>%
  ggplot() +
  geom_line(aes(year, value, color = asset, linetype = industry)) +
  scale_x_continuous(name = "", breaks = seq(1982, 2022, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

paperdata$asset %>%
  ggplot() +
  geom_line(aes(year, value_norm, color = asset, linetype = industry)) +
  scale_x_continuous(name = "", breaks = seq(1982, 2022, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  theme(legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1))

```

As with labor, most of the agricultural machinery is owned by farms.
Until 2002, the stock of tractors held by farmers and service providers was falling, as was the stock of other machinery owned by farmers.
Service providers were increasing their stock of other machinery in the 1980s, but by 2002 it returned to the level of 1982.
After 2002, farms start growing their capital and it returns to the levels of 1982 at about \$50B in tractors and \$180B in other machinery.
Service providers continue getting rid of tractors, falling down to only \$0.5B in 2022 compared to \$10.9B in 1982.
However, they begin to rapidly invest in other agricultural machinery, going from \$5.4B in 2002 to \$38.9B - a remarkable 7-fold increase over only 20 years.
The difference in this trend compared to that of tractors - a more general purpose machinery - is consistent with the narrative of service providers focusing on more specialized processes required a more specialized type of asset.

We summarize the trends in the management of labor and capital by farms and service providers in [@fig-asset-sh].
Tractors and other agricultural machinery are combined into a single category.
Since 1982, the share of agricultural machinery owned by service providers went up from 6.1% to 14.1%, and the share of paid labor hired through intermediaries increased from 11.6% to 19.5%.
The trends were not always monotone, and the providers' share of machinery was in fact falling in the 1990s, but these figures emphasize the significant role that the services sector plays in managing agricultural production inputs, especially remembering that the number of businesses in this industry is only about 1% of the number of crop farms.

```{r}
#| label: fig-asset-sh
#| fig-cap: "Share of capital and labor managed by service providers"

# paperdata$asset_sh %>%
#   filter(year %in% c(1982, 2022))

paperdata$asset_sh %>%
  ggplot() +
  geom_line(aes(year, share, linetype = name)) +
  scale_x_continuous(name = "", breaks = seq(1982, 2022, 10)) +
  scale_y_continuous(name = "", limits = c(0, 0.25), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  theme(legend.title = element_blank(), legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1)) 
```

The main goal of this paper is to identify how vertical disintegration of the agricultural production that manifests in the increasing share of farm activities performed by non-farm businesses affects agricultural productivity, which has also shown spectacular growth for over than 70 years of measurement in the United States.
Out of 36 industries, agricultural productivity growth was outpaced only in the Computer and electronic products sector [@eldridge2020].
Admittedly, common trend in productivity and use of services is a weak form of evidence, and a lot of other factors such as agricultural production technology, quality of inputs and scale of operations have also evolved in parallel.

The golden standard of productivity measurement is the total factor productivity (TFP) growth accounting, methodology pioneered by @solow1957 and further refined by @jorgenson1967.
Under this approach, TFP growth is measured as the difference - or residual - between the growth rates of output and input, adjusted for quality changes.
Quality adjustment plays a significant role in growth accounting for many industries and prevents misattributing improvements in the quality of inputs - such as more efficient pesticides, faster computers or better educated workerforce - to improvements in the production technology.

In agriculture, farm output nearly tripled since 1948 while use of quality adjusted farm inputs has remained flat, implying an annualized productivity growth rate of 1.49 percent [@usda-ers2025; @wang2024].
This growth has mainly been attributed to technological improvement, primarily fueled by R&D investments from public and private sectors, enhanced by public infrastructure, extension, and technology spillover from other sectors or neighboring regions [@alston2010; @huffman2006; @fuglie2007; @plastina2012; @wang2013].
However, it is very hard to assess the role that the growing services sector might have played in the evolution of agricultural productivity due to the data limitations and the assumptions imposed by this macro level approach.

In this study, we estimate the relationship between farm productivity and agricultural support services using both cross-sectional and time-series variation at a county level between 2002 and 2022.
TFP (multi-factor) estimates are not available at this geographic scale, and instead we turn to a simpler and widely used land (single-factor) productivity, measured in units of output per one acre of harvested cropland.
Output can be either measured in dollars of sales, or in physical units such as bushels.
Akin to the TFP, the former measure allows to aggregate all crop types and include such operational decisions as crop composition.
The latter measure, commonly known as yield, captures a more narrow definition of productivity and focuses on specific crops.

@tbl-desc-yield shows summary statistics of sales per acre and yields of the three crops - corn, soybeans and wheat - used in this paper in the sample of counties with active presence of agricultural service providers.
Notably, the number of county observations in each crop category is smaller than the number of counties with respective farm production, reflecting the fact that not all counties have a service provider available.
Both mean and median land productivity is increasing for all crop categories in almost all 5-year intervals and is widely distributed across counties.

```{r}
#| label: tbl-desc-yield
#| tbl-cap: "Land productivity in sales (all crops) or bushels (corn, soybeans and wheat) per acre."

paperdata$desc_yield

# figure for easier visual read
# paperdata$desc_yield %>%
#   ggplot(aes(x = year)) +
#   geom_boxplot(aes(lower = `25%`, upper = `75%`, middle = `50%`, ymin = `25%`, ymax = `75%`, group = year), stat = "identity", width = 2) +
#   geom_line(aes(y = mean)) +
#   facet_wrap("commodity", scales = "free_y")
```

# 3. Methodology

*This section will be further expanded.*

<!--# TODO:Verify in the lab:- are regressions weighted? -->

## Data sources

-   Census of Agriculture, 2002-2022.
    Definition of county samples.
    Imputation of missing values.
    Approximation of land HHI.

-   Quarterly Census of Employment and Wages (QCEW).
    Publicly available NAICS 1151 data, subject to significant suppression at a county level.

-   Longitudinal Business Database, 2022-2017.
    Establishment level microdata constructed from administrative IRS tax records.
    Aggregated to county level observations.

## Regression models

Productivity, payroll, land and average land are always log-transformed.
Productivity is either sales per acre or yield (bushels per acre).

Static productivity without interaction:

$$
Productivity = Payroll + Land + AvgLand + LandHHI 
$$ {#eq-reg-static}

Static productivity with interaction:

$$
Productivity = Payroll + Land + Payroll * Land + AvgLand + Payroll * AvgLand + LandHHI 
$$ {#eq-reg-static-int}

Dynamic productivity, $\Delta Productivity(t) = Productivity(t+5) - Productivity(t)$ in logs:

$$
\Delta Productivity = Productivity + Payroll + Land + AvgLand + LandHHI 
$$ {#eq-reg-dynamic}

# 4. Results

A key contribution of our study is the use of the Longitudinal Business Database (LBD) to measure the activity of the agricultural services providers in NAICS industry 1151.
This dataset is only accessible through an approved research project in the FSRDC environment, but the best publicly available data source is vastly inadequate to perform the analysis of the relationship between farm and services activity.
@tbl-qcew-sup shows aggregation of county level QCEW data on NAICS 1151 establishments, employment and payroll merged with Census of Agriculture measures of harvested land, sales and the number of farms engaged in crop production in 2012.
Counties included in the "suppressed" column only have the number of service provider establishments reported, while employment and payroll are suppressed to protect business data confidentiality.
The 76% of counties with suppression host 38% of all agricultural services establishments.
And while these establishments are mostly small and constitute only 15% of total employment and 18% of total payroll, they operate in counties where more than half of all crop production is taking place.

```{r}
#| label: tbl-qcew-sup
#| tbl-cap: Data suppression in 2012 QCEW.

paperdata$qcew_sup %>%
  relocate(name, total, no_agser, published, suppressed, sh_suppressed) %>%
  rename(Measure = name, Total = total, `No services` = no_agser, Available = published, Suppressed = suppressed, `% suppr.` = sh_suppressed) %>%
  tt() %>%
  format_tt(i = 1, j = 2:5, digits = 0, num_mark_big = ",") %>%
  format_tt(i = 2:7, j = 2:5, digits = 3, num_suffix = TRUE) %>%
  format_tt(j = 6, fn = scales::label_percent()) %>%
  style_tt(j = 1, align = "l")
```

Using the LBD, we can characterize the structure of the local agricultural services market.
@tbl-agser-distr reports the evolution of the number of establishments, payroll and market concentration in the NAICS 1151 sector at the county level from 2002 to 2017.
In line with the national aggregate trends reported in @sec-background, the average number of establishments has been slowly declining while the average real payroll has nearly doubled from \$2.4 to \$4.0 million.
Notably, there is no evidence of an increasing local market concentration, as the mean and median payroll shares of the largest establishment in a county remained fairly stable around 73% and 78% respectively.
This finding is in sharp contrast to the trends of increasing market concentration in the agricultural production on the farm and in the adjacent markets along the production chain beyond the farm gate.
<!--# TODO: elaborate more on this, add farm distribution numbers and quote findings from the market structure literature -->

```{r}
#| label: tbl-agser-distr
#| tbl-cap: "County distribution of service providers"

paperdata$agser_distr %>%
  select(variable, year, nobs, mean, sd, `25%`, `50%`, `75%`) %>%
  arrange(variable, year) %>%
  mutate(
    variable = case_match(
      variable,
      "est" ~ "Establishments",
      "pay" ~ "Payroll",
      "pay_cr1" ~ "Largest payroll"
    )) %>%
  tt() %>%
  format_tt(j = 3, digits = 0, num_mark_big = ",") %>%
  format_tt(i = 1:4, j = 4:8, digits = 3) %>%
  format_tt(i = 5:8, j = 4:8, digits = 3, num_suffix = TRUE) %>%
  format_tt(i = 9:12, j = 4:8, fn = scales::label_percent()) %>%
  style_tt(i = 1, j = 1, rowspan = 4) %>%
  style_tt(i = 5, j = 1, rowspan = 4) %>%
  style_tt(i = 9, j = 1, rowspan = 4)

```

Regression results for the estimation of the model in [@eq-reg-static-int] for all-crop sample in 2017 are presented in [@tbl-reg-static-crop].
Coefficients of lesser interest have been suppressed for confidentiality disclosure avoidance, showing only the sign and statistical significance.
This model shows there is a positive and statistically significant relationship between farm crop productivity and total payroll of the service providers operating in a county, and this effect persist controlling for land distribution variables.
A 1% increase in the payroll size of the services sector is associated with a 0.2% increase in crop sales per acre of harvested land.
We interpret this as an evidence of the positive impact of availability of service providers on farm productivity.
The effect is stronger in counties with more total acreage and smaller average farm size, consistent with a story of service providers becoming more effective when serving many small farms at scale.

```{r}
#| label: tbl-reg-static-crop
#| tbl-cap: "Regression of crop sales per acre on service provider payroll, 2017."

models <- paperdata$reg_static %>%
  filter(commodity == "crop", year == 2017)

fits <- unique(models$model) %>%
  map(\(l) filter(models, model == l))

modelsummary(fits, output = "huxtable")
```

The advantage of using dollar crop sales per acre as the measure of agricultural land productivity is the ability to aggregate all crop production into a single number.
But it comes at a cost of obscuring the variation of productivity and its responsiveness to availability of service providers across different crop categories.
In particular, vegetables, fruits and horticulture tend to return higher sales per acre than grains while also being more labor intensive to grow.
Since our measure of service provider size - payroll to employees - most directly reflects the labor component of support activity, counties with sales dominated by vegetables, fruits and horticulture are more likely to have service providers with high payrolls regardless of farm productivity, leading to an upward bias in the estimated coefficient.
As we do not observe in the data which crops service providers are involved with, our approach to address this issue is to look at yields of three major homogeneous grain crops - corn, soybeans and wheat - measured in bushels per acre and to restrict the samples to counties with positive production of respective crops.
In a symmetric fashion, this approach would attenuate the coefficient of interest because the payroll is potentially affecting productivity of commodities other than the one which yield is being used as the outcome in the model.
Measuring output in bushels instead of dollars also avoids the issue of unobserved unit price variation.

[@tbl-reg-static-com] presents payroll coefficients for the static productivity model without interaction estimated over corn, soybeans and wheat samples in 2002, 2007, 2012 and 2017.
A 1% increase in agricultural service payroll is associated with a 0.02-0.06, 0.01-0.03 and 0.02-0.05 percent increase in corn, soybeans and wheat yields, respectively.
While the magnitude of these estimates is smaller than in the sales-per-acre model, they are always statistically significant and likely underestimated as discussed above.

```{r}
#| label: tbl-reg-static-com
#| tbl-cap: "Payroll coefficient estimates"

models <- paperdata$reg_static %>%
  filter(model == "logyield ~ logpay + logland + logavgland + hhi_land",
         term == "logpay",
         commodity %in% c("corn", "soy", "wheat"))

models %>%
  mutate(value = paste0(round(estimate, 3), significance, " (", round(std.error, 3), ")")) %>%
  pivot_wider(id_cols = year, names_from = commodity) %>%
  datasummary_df(fmt = 0)


```

The effect of the service payroll on yields varies by the distribution of harvested land, as identified by the interaction terms in [@eq-reg-static-int].
Full set of estimates are reported in [Tables -@tbl-reg-int-1], [-@tbl-reg-int-2] and [-@tbl-reg-int-3], and for easier interpretation, [@fig-me] shows marginal effects of payroll evaluated at high and low quartiles of the respective land variables.
Large magnitudes of variation of the marginal effect across time and commodities suggest that heterogeneity might be present, but statistical significance of the estimates is generally low, and no strong patterns emerge.

```{r}
#| label: tbl-reg-int-1
#| tbl-cap: "Regression of corn yields on service provider payroll."

coef_renames <- c(
  "logpay" = "Payroll", 
  "logpay:logland" = "Payroll*Land", 
  "logpay:logavgland" = "Payroll*AvgLand", 
  "logland" = "Land", 
  "logavgland" = "AvgLand", 
  "hhi_land" = "LandHHI")

paperdata$reg_int %>%
  filter(commodity == "corn") %>%
  pull(model, name = year) %>%
  modelsummary(output = "huxtable", coef_map = coef_renames)
```

```{r}
#| label: tbl-reg-int-2
#| tbl-cap: "Regression of soybean yields on service provider payroll."

paperdata$reg_int %>%
  filter(commodity == "soy") %>%
  pull(model, name = year) %>%
  modelsummary(output = "huxtable", coef_map = coef_renames)
```

```{r}
#| label: tbl-reg-int-3
#| tbl-cap: "Regression of wheat yields on service provider payroll."
 
paperdata$reg_int %>%
  filter(commodity == "wheat") %>%
  pull(model, name = year) %>%
  modelsummary(output = "huxtable", coef_map = coef_renames)
```

```{r}
#| label: fig-me
#| fig-cap: "Marginal effect of payroll on yield"
#| fig-subcap:
#|   - "At different levels of land"
#|   - "At different levels of average land"
#| layout-nrow: 2
#| fig-asp: 0.8

paperdata$reg_me %>%
  filter(atvar == "land") %>%
  ggplot() +
  geom_ribbon(aes(logland, ymin = lo, ymax = hi), alpha = 0.2) +
  geom_line(aes(logland, dydx)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  facet_wrap(c("commodity", "year"), scales = "free_y")

paperdata$reg_me %>%
  filter(atvar == "avgland") %>%
  ggplot() +
  geom_ribbon(aes(logavgland, ymin = lo, ymax = hi), alpha = 0.2) +
  geom_line(aes(logavgland, dydx)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  facet_wrap(c("commodity", "year"), scales = "free_y")

```

We recognize that the cross-sectional estimates presented here are possibly subject to endogeneity bias.
While productivity enhancing effect of agricultural service provider availability is a reasonable explanation for the observed positive relationship, the effect might go in the opposite or, more likely, in both directions: The most productive farms are more dynamic and innovative, and create demand for specialized service providers to appear in their local markets.
In this study, we begin to untangle this relationship by analysing the dynamic response of farm yields to provider availability.
Model presented in [@eq-reg-dynamic] evaluates whether counties with more agricultural provider activity demonstrated higher growth rate in yields over the subsequent five year period.
<!--# TODO: explain why we control for base year productivity in these regressions --> Payroll coefficient estimates are collected in [@tbl-reg-dynamic].
All statistically significant coefficients in these models are positive, supporting the hypothesis that agricultural service providers increase farm productivity.
To illustrate with the largest coefficient, crop-growing counties that had a 1% larger service sector experienced a 0.05% growth rate in corn yields.

```{r}
#| label: tbl-reg-dynamic
#| tbl-cap: Payroll coefficient estimates

models <- paperdata$reg_dynamic %>%
  filter(model == "diff_logyield ~ logpay + logyield + logland + logavgland + hhi_land",
         term == "logpay")

models %>%
  mutate(value = paste0(round(estimate, 3), replace_na(significance, ""), " (", round(std.error, 3), ")")) %>%
  pivot_wider(id_cols = year, names_from = commodity) %>%
  datasummary_df(fmt = 0)


```

# 5. Discussion

In this paper, we present new estimates of the relationship between farm productivity and agricultural support services using both cross-sectional and time-series variation.

Results using publicly available data from the Quarterly Census of Employment and Wages (QCEW) indicate that in the cross section, greater access to agricultural service providers is associated with higher farm productivity.
This relationship appears to be highly heterogeneous, varying with the agricultural commodity and the service provided.
It also depends on the size of farm operations.
However, a noteworthy limitation of using publicly available information on service providers is the high levels of data suppression that occur to maintain disclosure avoidance standards.
Particularly when using county-level data at the six-digit NAICS industry level, the plurality of observations is suppressed.
Further, the treatment of agricultural and non-agricultural labor varies across statistical agency, data program, and administrative purpose.
Therefore, a significant innovation of this study is the use of the unsuppressed Federal administrative data on establishment payroll tax filings from the Internal Revenue Service contained in the Longitudinal Business Database for businesses in NAICS industry 1151 (Support Activities for Crop Production).

Using the LBD and the Census of Agriculture for 2002, 2007, 2012 and 2017, we find that 1% larger agricultural support services sector in a county, as measured by total payroll, is associated with 0.2% higher farm sales per acre, controlling for farm characteristics.
The relationship is also strong for corn and wheat yields, but not for soybean yields.
Positive effect on productivity is amplified in counties with larger farmland acreage.
Leveraging the longitudinal nature of the LBD, we find that counties with higher service provider availability experience higher productivity growth over the subsequent 5-year period.

## Implications for productivity accounting

TODO: revise to fit the concluding section

Decomposition of farm input usage reveals that large reductions in use of land and labor have been offset by increases in capital and intermediate inputs, the latter showing an increase of 109 percent from 1948 to 2021 [@wang2024].
Just as economies of scale may accrue to larger farms, support service providers may be able to increase capital intensity and adopt new technologies more quickly.
By focusing on a narrowing set of production activities, service providers may also experience significant gains from specialization.
It might well be that existing quality adjustments to the cost of intermediate inputs are underestimated.
Specifically, returns to scale and specialization may make agricultural service providers more productive than farmers at on-farm production activities, increasing outputs while decreasing costs.

A key component of modern agricultural productivity accounting is quality adjustment.
TFP growth is calculated as the difference between the growth rates of output and inputs, and aggregation across different commodity and expenditure types necessitates the use of dollar denominated nominal quantities, which are in turn deflated to real terms using a combination of price indices.
The price index is constructed to account both for relative price variation over time and geography for factors where quality does not change (e.g. energy) or quality measurement is not available (farm products), and for changes in quality coming from either intrinsic quality improvements (a pesticide becoming more potent) or substitution between different types (different education attainment of workers).
The classic argument of @jorgenson1967 is that the estimate of TFP or “the measure of our ignorance” can be explained away by accurately accounting for changes in all the inputs, and by doing so we can better understand the sources rather than just the magnitude of growth.
Indeed, input quality changes in labor, capital, and intermediate inputs explain 0.11, 0.04, and 0.04 percentage points, respectively, of the 1.46 percentage points of annual agricultural output growth [@wang2024, p. 26].

Depending on data availability, @wang2024 use two methods to calculate quality-adjusted price indices – hedonic and index number approach.
The index number method requires detailed and disaggregated data on prices, quantities and composition of characteristics of a particular input.
The hedonic method applies regression analysis to estimate year dummy coefficients and can be used with less available data.
While both methods allow for a more accurate measurement of input volumes in constant efficiency units, it is important to recognize their limitations that may lead to biases.

Two primary farm expenditure categories associated with agricultural service providers are contract labor and purchased machine services.
Nominal contract labor expenditures are deflated with a hedonic price index of agricultural worker earnings, estimated from the National Agricultural Workers Survey (NAWS) that includes an indicator variable of whether the worker was employed by a contractor [@wang2025].
This approach relies on the assumptions of representative sampling and correct model specification, with potential biases for price index estimation in both directions.

Purchased machine services are considered a substitute for own-capital input, and the same price index is applied for quality adjustment of both expenditure types.
Positive relationship between the use of service providers and crop yields established in this paper implies that machinery managed by providers is used more productively and offers more constant efficiency units of input for a given level of nominal spending.
Not accounting for these differences in machinery input quality leads to underestimation in the growth rate of quality-adjusted service input and overestimation of farm TFP growth.

Our findings on the heterogeneity of the yield-provider relationship across crop types and land acreages suggest an avenue for a more accurate quality adjustment of service expenditures.
Similarly to how labor expenditures are broken down into demographic groups by worker characteristics (gender, age, education and employment class), each with its own productivity and growth rates, service expenditures can be also broken down based on the characteristics of the farms that employ them.
Service expenditures in groups with little effect of providers on productivity, such as soybean farms, should be adjusted using the same price index as for owned machinery, and with lower price index (i.e. higher constant efficiency quantity) in groups with higher effect, such as wheat and corn farms.

While the national productivity accounting framework can be used to study the effect of agricultural service providers on the aggregate farm productivity, the accuracy of this approach would only be as good as the accuracy of the quality adjustment procedures, which currently suffers from the aforementioned limitations.
In this paper, we contribute to the agricultural productivity discussion by generating estimates of the relationship between farm productivity and scale of provider operation identified from the spatial variation across United States counties.
This approach allows us to draw further insights on the heterogeneity of this relationship over different crop types and farm land distributions.

The theoretical foundation of productivity accounting also relies on the assumption of perfectly competitive market equilibria where output prices are equal to marginal costs of production, and input factors are paid their respective marginal product (see [@hulten2010, p. 1023] for a brief overview).
Under this assumption, cost shares are equal to the corresponding output elasticities and can thus be used in the construction of the Divisia-Tornkvist growth indices.
In practice, however, product and input markets are often subject to varying degrees of market power on either side of the trade, introducing distortions to the calculation of quality adjusted quantities.
For illustration, if agricultural service providers are increasing their trading power in the local markets over time, service price will be increasing over its marginal product, but for an econometrician estimating a hedonic price index this increase will be indistinguishable from an increase in the unobserved quality of service, leading to biases in the estimates of constant efficiency quantity and TFP.

Agricultural service providers can be thought of as intermediaries between farms and agricultural inputs.
This perspective helps think about implications of market power for TFP growth accounting.
For example, if farms hire labor directly, individual farm's market power is likely low, and workers are paid their marginal product.
Then labor share of farm revenue correctly estimates output elasticity of labor.
But, if introducing an intermediation layer creates a degree of monopsony power of service providers over workers, they are paid less than their marginal product, and the income share of contract labor expenditures underestimates its output elasticity, leading to overestimation of TFP.
Market power over contract labor is a likely possibility given market concentration of service providers in local county markets is often very high as we show in this paper and contract workers are paid less than hired workers of same qualifications [@wang2025].
The resulting overestimation of farm TFP growth becomes more important as the share of contract labor in farm labor expenditures is increasing.

# References

::: {#refs}
:::

# Cuts from old versions

## Introduction

*LBD is a feasible data source to study dynamism of the services sector.*

Despite the importance of the agricultural services sector to the farm economy, because of gaps in the federal statistical system, relatively little is known about how this industry has evolved over the past three decades.
Using the Longitudinal Business Database (LBD) – a restricted use dataset constructed from the administrative federal tax records – @babkin2022 document establishment, employment, payroll, and wage dynamics for businesses that provide support services to crop production (NAICS industry code 1151).
They demonstrate that agricultural support services is a uniquely dynamic and rapidly evolving sector of the American economy.
Although rates of entry have fallen substantially, large incumbent firms have tended to grow in size.
Furthermore, the real annual mean wage to nonagricultural workers employed in the sector has increased by nearly \$10,000 over the 1992-2016 period, indicating sizable growth in labor productivity.
Nonetheless, standard measures of economic dynamism, a principal source of productivity growth, have all fallen, which may have implications for innovation and the long-term health of this industry and agricultural production more generally.

*Agricultural productivity growth is a big deal, and this is how it is measured at macro level.*

One of the most robust stylized facts about the agricultural production is the secular increase in farm total factor productivity (TFP).
In the aggregate productivity accounting, pioneered by @solow1957 and further refined by @jorgenson1967, TFP is measured as the residual from the growth rate of output net of the growth rates of inputs, adjusted for changes in quality.
Since 1948, farm output nearly tripled while use of farm inputs has remained flat, implying an annualized productivity growth rate of 1.49 percent [@usda-ers2025; @wang2024].
This is remarkable, as out of 36 industries, this growth rate was only outpaced by the Computer and electronic products sector [@eldridge2020].
Agricultural productivity growth has mainly been attributed to technological improvement, primarily fueled by R&D investments from public and private sectors, enhanced by public infrastructure, extension, and technology spillover from other sectors or neighboring regions [@alston2010; @huffman2006; @fuglie2007; @plastina2012; @wang2013].
However, it is very hard to assess the role that the growing services sector might have played in the evolution of agricultural productivity due to the data limitations and the assumptions imposed by this macro level approach.

*Another approach is to productivity measurement is the micro-econometric perspective.*

*Reallocation is where micro meets macro. Labor and capital move from farms to services.*

The aggregate productivity accounting approach based on the original Solow–Jorgenson–Griliches paradigm has been greatly enriched by the econometric analysis of longitudinal establishment-level datasets [@bartelsman2000; @foster2001; @foster2008].
The panel nature of these data provides information on business entry and exit, allowing to separate the productivity changes within a firm from the aggregate growth resulting from reallocation of resources between firms.

The classic "creative destruction" argument maintains that the reallocation from less to more effective users is a major factor in the increase of the aggregate productivity.
The observed shift of labor, capital and managerial resources from farms to specialized service providers can be seen through this lens as a reallocation to businesses that can utilize them in a more efficient manner, implying both an increase in the aggregate productivity of the extended agricultural sector and a higher average productivity of the remaining farms through positive selection.

*Inter-sectoral reallocation has been studied in the outsourcing literature. The productivity implications are ambiguous.*

Outside of the agricultural sector, the effects of vertical desintegration on firm performance have been studied in the context of outsourcing in manufacturing and services.
It has been shown both theoretically and empirically that outsourcing is not always associated with increases in productivity [@heshmati2003; @görg2004; @lahiri2016].
Outsourcing is expected to involve production cost savings relative to internal production because outside suppliers benefit from economies of scale, smoother production schedules and centralization of expertise, but the choice between continued internal production or an outsourcing decision requires other considerations than pure production cost differences.

Theoretical literature suggests that the outsourcing decision can be viewed through multiple lenses.
Transaction cost economics argues that the optimal form of organization is primarily a function of the characteristics underlying a given exchange between a provider and its client, and that unfavorable exchange conditions can increase the cost of writing enforceable contracts [@coase1937; @williamson2010].
According to the resource-based view of the firm, outsourcing allows clients to gain access to specialized resources and capabilities residing in the provider firms and to focus more on core competencies [@barney1991].
From the perspective of the agency theory, client firms (principals) employ providers (agents) to perform certain functions or processes on their behalf, and the expected returns can be lowered by the risks of opportunism and moral hazard [@ross1973].
Multiple other theoretical approaches exist, but regardless of model predictions, realized effects of outsourcing on firm performance depend greatly on the implementation of the outsourcing strategy by firm management.

Empirically, a range of studies from different national contexts have shown that outsourcing can have positive [@bardhan2006], negative [@broedner2009] or no significant effects [@bengtsson2009] on firm performance.
In is worth noting that empirical studies use a wide array of performance indicators, including financial metrics (return on assets, profit margins, sales), labor productivity and indirect measures such as customer satisfaction or rates of innovation.

*We turn to dis-aggregated data (county and establishment level) to study the reallocation of management, capital and labor from farms to agser providers.*

In this study, we measure the scale of activity of the agricultural service providers in a county using total payroll of the LBD establishments in the NAICS 1151 industry.
These data offer multiple advantage over the alternative of using expenses reported in the Census of Agriculture (COA).
First, COA is a survey over a sample of farms with an overall response rate of 61% in the 2022 round [@usda-nass2024] and lower rates for individual questions.
LBD, on the other hand, is based on administrative IRS records and includes every active establishment with positive payroll, providing universal coverage and highest data accuracy.
Second, payroll expenditures paid by service providers to workers better track the quantity of labor input used in production than the amounts paid to providers by farmer.
Finally, the use of microdata resolves the issue of cell suppression and allows us to calculate custom indicators characterizing local markets, such as measures of market concentration.

*Aggregated data from the Census of Agriculture do not allow separation of farm expenditures by farm output, and thus it is not possible to tell what share of expenditures goes to crop and animal production, whereas service provider establishments can be easily separated according to the NAICS code.*
