---
title: "2025 AAEA conference - Ag services"
format:
  html:
    date: today
    toc: true
    toc-depth: 2
    embed-resources: true
    page-layout: full
    code-fold: true
    df-print: paged
---

# top

All dollars values are in real terms.

```{r}
#| output: false

library(logger)
library(tidyverse)
library(tmap)
library(tigris)
library(sf)
library(skimr)
library(modelsummary)
library(estimatr)
library(patchwork)

theme_set(theme_minimal())

log_threshold(WARN)
options(tigris_use_cache = TRUE)


# scripts
source("projects/aaea-conf-2025-agser/agser.R", local = (agser <- new.env()))
glance.lmdf <- agser$glance.lmdf
tidy.lmdf <- agser$tidy.lmdf


PAR <- list(
  lbd_years = c(2002, 2007, 2012, 2017),
  commodities = c("crop", "corn", "soy", "wheat")
)



venn <- function(x, y) {
  and <- base::intersect(x, y)
  or <- base::union(x, y)
  list(
    x = x,
    y = y,
    and = and,
    or = or,
    x_y = base::setdiff(x, y),
    y_x = base::setdiff(y, x),
    xor = base::setdiff(or, and)
  )
}

```

## data

```{r}
dat <- list(
  tfp = agser$data_tfp(),
  farm_nat = agser$data_farm("national"),
  farm_cty = agser$data_farm("county"),
  farm_labor_pre2002 = agser$data_farm_labor_pre2002(),
  agser_nat = agser$agser_nat(),
  io = agser$data_io(),
  fa = agser$data_fa(),
  rdc_2019 = agser$rdc_res_2019(),
  rdc_2024 = agser$rdc_res_2024()
)

paperdata <- list()

```


```{r}
#| eval: false
saveRDS(paperdata, agser$opath$paper)
```


# National farm productivity

## TFP

### 1948 = 100%

Reproduce classic ERS [figures](https://www.ers.usda.gov/data-products/agricultural-productivity-in-the-united-states/summary-of-recent-findings) for verification.

```{r}
d <- dat$tfp %>%
  select(year, output, input, tfp, capital, labor, starts_with("int")) %>%
  mutate(across(!year, \(x) x / first(x))) %>%
  pivot_longer(!year)

# tfp
d %>%
  filter(name %in% c("output", "input", "tfp")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name))

# input decomposition
d %>%
  filter(name %in% c("capital", "labor", "int")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name))

# intermediates
d %>%
  filter(str_starts(name, "int_")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name))

# intermediates, no pesticide
d %>%
  filter(str_starts(name, "int_"), name != "int_pest") %>%
  ggplot() +
  geom_line(aes(year, value, color = name))

```

### 2002 = 100%


```{r}

paperdata$tfp <- d <- dat$tfp %>%
  select(year, output, input, tfp, capital, labor, starts_with("int")) %>%
  filter(year >= 2002) %>%
  mutate(across(!year, \(x) x / first(x))) %>%
  pivot_longer(!year)

d %>%
  filter(name %in% c("output", "input", "tfp")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name))

# input decomposition, 2002 = 100%
d %>%
  filter(name %in% c("capital", "labor", "int")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name))

# intermediates, 2002 = 100%
d %>%
  filter(str_starts(name, "int_")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name))

```

## Land productivity

Partial factor productivity.

Aggregate national statistics from Census of Agriculture and BEA.

Sales per acre give a simple measure that combines all crops, but are affected by price fluctuations and lump together row crops with fruits and vegetables.
Yields (production per acre) are crop-specific, and we look at corn, soybeans and wheat.


### sales per acre

Increase over time with nearly constant total acreage.

```{r}
#| fig-width: 10
#| fig-height: 4

d <- dat$farm_nat %>%
  filter(name %in% c("sale_crop", "land_crop")) %>%
  pivot_wider(id_cols = year) %>%
  mutate(sale_per_acre = sale_crop / land_crop) %>%
  pivot_longer(!year)

p1 <- d %>%
  filter(name == "sale_crop") %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  labs(title = "Crop sales")

p2 <- d %>%
  filter(name == "land_crop") %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  labs(title = "Harvested cropland")

p3 <- d %>%
  filter(name == "sale_per_acre") %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  labs(title = "Sales per acre")

p1 + p2 + p3
```


### yield by crop

Varying dynamic across crop types, but yields generally increase.

```{r}
d <- dat$farm_nat %>%
  filter(str_detect(name, "^(land|prod)_(corn|soy|wheat)")) %>%
  separate_wider_delim(name, "_", names = c("name", "com")) %>%
  pivot_wider(id_cols = c(year, com)) %>%
  mutate(yield = prod / land) %>%
  pivot_longer(!c(year, com)) %>%
  mutate(name = ordered(name, c("prod", "land", "yield")))

d %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  facet_wrap(vars(com, name), ncol = 3, scales = "free_y") +
  labs(title = "Crop land, production and yield")

d %>%
  arrange(com, name, year) %>%
  mutate(value = value / first(value), .by = c(com, name)) %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0.8, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  facet_wrap(vars(com, name), ncol = 3, scales = "fixed") +
  labs(title = "Crop land, production and yield, 2002 = 100%")

```


### agser sample stats

2000 counties: not all counties with farming have service providers.

```{r}
paperdata$desc_yield <- d <- dat$rdc_2024$desc %>%
  filter(variable == "yield") %>%
  select(commodity, year, nobs, mean, sd, "25%", "50%", "75%") %>%
  mutate(commodity = case_match(commodity, "crop" ~ "all crops", "soy" ~ "soybeans", .default = commodity)) %>%
  arrange(commodity, year)
d

d %>%
  ggplot(aes(x = year)) +
  geom_boxplot(aes(lower = `25%`, upper = `75%`, middle = `50%`, ymin = `25%`, ymax = `75%`, group = year), stat = "identity", width = 2) +
  geom_line(aes(y = mean)) +
  facet_wrap("commodity", scales = "free_y")
```



# Farm expenses

## Ag census

Multiple categories relate to ag services.
Contract labor expense also goes to providers.

```{r}
#| rows.print: 20
#| column: screen-inset
tribble(
  ~name,            ~short_desc,
  "exp_tot",         "EXPENSE TOTALS, OPERATING - EXPENSE, MEASURED IN $",
  "exp_fert",        "FERTILIZER TOTALS, INCL LIME & SOIL CONDITIONERS - EXPENSE, MEASURED IN $",
  "exp_chem",        "CHEMICAL TOTALS - EXPENSE, MEASURED IN $",
  "exp_seed",        "SEEDS & PLANTS TOTALS - EXPENSE, MEASURED IN $",
  "exp_anim",        "ANIMAL TOTALS - EXPENSE, MEASURED IN $", # Livestock and poultry purchased or leased
  "exp_feed",        "FEED - EXPENSE, MEASURED IN $",
  "exp_fuel",        "FUELS, INCL LUBRICANTS - EXPENSE, MEASURED IN $", # Gasoline, fuels, and oils purchased
  "exp_rent",        "RENT, CASH, LAND & BUILDINGS - EXPENSE, MEASURED IN $",
  "exp_repair",      "SUPPLIES & REPAIRS, (EXCL LUBRICANTS) - EXPENSE, MEASURED IN $",
  "exp_labhire",     "LABOR, HIRED - EXPENSE, MEASURED IN $",
  "exp_labcont",     "LABOR, CONTRACT - EXPENSE, MEASURED IN $",
  "exp_agser_cust",  "AG SERVICES, CUSTOMWORK - EXPENSE, MEASURED IN $",
  "exp_agser_util",  "AG SERVICES, UTILITIES - EXPENSE, MEASURED IN $",
  "exp_agser_rent",  "AG SERVICES, MACHINERY RENTAL - EXPENSE, MEASURED IN $",
  "exp_agser_anim",  "AG SERVICES, CUSTOM SERVICES FOR LIVESTOCK, INCL MEDICAL SUPPLIES & VETERINARY - EXPENSE, MEASURED IN $",
  "exp_agser_other", "AG SERVICES, OTHER - EXPENSE, MEASURED IN $",
  "exp_int",         "INTEREST - EXPENSE, MEASURED IN $",
  "exp_tax",         "TAXES, PROPERTY, REAL ESTATE & NON-REAL ESTATE, (EXCL PAID BY LANDLORD) - EXPENSE, MEASURED IN $"
  ) %>%
  mutate(short_desc = str_sub(short_desc, 1, -25))
```

- All crop related expenses have been growing over time.
- Ag services - at a very slow rate, +25% over 20 years.
- Contract labor expenses - one of the fastest growths, nearly doubled.

```{r}
d <- dat$farm_nat %>%
  filter(str_starts(name, "exp_"), !(name %in% c("exp_tot", "exp_anim", "exp_feed", "exp_int", "exp_tax", "exp_agser_anim"))) %>%
  mutate(name = if_else(str_detect(name, "agser"), "exp_agser", name)) %>%
  summarize(value = sum(value), .by = c(year, name))

d %>%
  ggplot() +
  geom_line(aes(year, value)) + 
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  facet_wrap("name", scales = "free_y") +
  labs(title = "Crop related expenses")

d %>%
  arrange(name, year) %>%
  mutate(value = value / first(value), .by = name) %>%
  ggplot() +
  geom_line(aes(year, value)) + 
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0.8, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  facet_wrap("name", scales = "fixed") +
  labs(title = "Crop related expenses, 2002 = 100%")

```




```{r}
d <- dat$farm_nat %>%
  filter(str_starts(name, "exp_agser"), name != "exp_agser_anim")

d %>%
  ggplot() +
  geom_line(aes(year, value)) + 
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  facet_wrap("name", scales = "free_y") +
  labs(title = "Ag services expenses")

d %>%
  arrange(name, year) %>%
  mutate(value = value / first(value), .by = name) %>%
  ggplot() +
  geom_line(aes(year, value)) + 
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0.8, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  facet_wrap("name", scales = "fixed") +
  labs(title = "Ag services expenses, 2002 = 100%")

```

### crop services and QCEW pay

```{r}
paperdata$exp_agser <- d <- bind_rows(
  dat$farm_nat %>%
    filter(name %in% c("exp_labcont", "exp_agser_cust")),
  dat$agser_nat %>%
    filter(between(year, 2002, 2022)) %>%
    mutate(name = "QCEW pay") %>%
    select(year, name, value = pay)) %>%
  select(year, name, value) %>%
  arrange(name, year)
d %>%
  ggplot() +
  geom_line(aes(year, value, linetype = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale()))
```


### labor

```{r}
paperdata$exp_labor <- d <- dat$farm_nat %>%
  filter(name %in% c("exp_labhire", "exp_labcont")) %>%
  bind_rows(dat$farm_labor_pre2002) %>%
  select(year, name, value) %>%
  arrange(name, year) %>%
  mutate(value_norm = value / first(value), value_gr = value / lag(value) - 1, .by = name)
d %>%
  ggplot() +
  geom_line(aes(year, value, linetype = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale()))
d %>%
  ggplot() +
  geom_line(aes(year, value_norm, linetype = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale()))
d %>%
  ggplot() +
  geom_line(aes(year, value_gr, linetype = name)) +
  scale_y_continuous(name = "", labels = scales::label_percent(scale_cut = scales::cut_short_scale()))

```


### county suppression

Custom work and hired labor expenses are suppressed in 100-200 counties every year.
This is not too bad - and much better than suppression of NAICS-1151 in QCEW.

```{r}
dat$farm_cty %>%
  filter(name %in% c("exp_labcont", "exp_agser_cust")) %>%
  count(name, year, value_f) %>%
  pivot_wider(names_from = value_f, values_from = n)
```


## BEA

Use of intermediate inputs and compensation of employees by crop industries.

```{r}
df_io <- dat$io %>%
  filter(com_code %in% c("115000", "T005", "V00100"), str_starts(ind_code, "111"))
df_io <- bind_rows(
  df_io,
  df_io %>%
    summarize(pur_val = sum(pur_val), .by = c(year, com_code, com_name)) %>%
    mutate(ind_code = "111", ind_name = "All crops")
) %>%
  mutate(
    com = case_match(
      com_code,
      "115000" ~ "agser",
      "T005" ~ "tot_int",
      "V00100" ~ "emp"
    ),
    ind = case_match(
      ind_code,
      "1111A0" ~ "oilseed",
      "1111B0" ~ "grain",
      "111200" ~ "vegetable, melon",
      "111300" ~ "fruit, nut",
      "111400" ~ "greenhouse",
      "111900" ~ "other",
      "111"    ~ "crops"
    )
  ) %>%
  rename(value = pur_val)

df_io %>%
  distinct(com, com_code, com_name)

df_io %>%
  distinct(ind, ind_code, ind_name)
```

- Ag service expenses grow faster than total intermediate expenses.
- Much faster than "compensation of employees" - not clear if it includes contract labor.

```{r}
#| fig-width: 10
#| fig-height: 4
d <- df_io %>%
  filter(ind == "crops") %>%
  mutate(com = ordered(com, c("agser", "tot_int", "emp")))

d %>%
  ggplot() +
  geom_line(aes(year, value)) + 
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  facet_wrap(c("com"), ncol = 3, scales = "free_y") +
  labs(title = "Crop industry expenses")

d %>%
  arrange(com, year) %>%
  mutate(value = value / first(value), .by = com) %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  facet_wrap("com", scales = "fixed") +
  labs(title = "Crop industry expenses, 2002 = 100%")

```

### agser expenses by crop

```{r}
#| fig-width: 10
#| fig-height: 6
d <- df_io %>%
  filter(ind != "crops") %>%
  mutate(ind = if_else(ind %in% c("grain", "oilseed"), ind, "all_other")) %>%
  summarize(value = sum(value), .by = c(year, ind, com)) %>%
  mutate(com = ordered(com, c("agser", "tot_int", "emp"))) %>%
  mutate(ind = ordered(ind, c("grain", "oilseed", "all_other")))

d %>%
  ggplot() +
  geom_line(aes(year, value)) + 
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  facet_wrap(c("ind", "com"), ncol = 3, scales = "free_y") +
  labs(title = "Expenses by crop")

d %>%
  arrange(ind, com, year) %>%
  mutate(value = value / first(value), .by = c(ind, com)) %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  facet_wrap(c("ind", "com"), ncol = 3, scales = "fixed") +
  labs(title = "Expenses by crop, 2002 = 100%")
```

# Farm capital

Stock of capital (machinery and equipment).
Did it move from farms to providers?

## Census


```{r}
#| fig-width: 10
#| fig-height: 4
d <- dat$farm_nat %>%
  filter(name == "asset_mach")

p1 <- d %>%
  ggplot() +
  geom_line(aes(year, value)) + 
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  labs(title = "Estimated market value of machinery and equipment")

p2 <- d %>%
  arrange(year) %>%
  mutate(value = value / first(value)) %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_x_continuous(name = "", breaks = unique(d$year)) +
  scale_y_continuous(name = "", limits = c(0.8, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  labs(title = "2002 = 100%")

p1 + p2
```

## BEA

BEA fixed asset accounts allow us to evaluate capital stock of ag service providers.
Although the industry is lumped with fishing forestry and fishing, we can look at equipment classes related to agriculture: "Farm tractors" and "Other agricultural machinery".

An interesting patterns stands out.
While farms roughly doubled stock of tractors and other machinery, providers stock of tractors was small and declining, but stock of other machinery increased tenfold.


```{r}
paperdata$asset <- d <- dat$fa %>%
  mutate(asset = case_match(asset_code, "EO30" ~ "other ag machine", "EO21" ~ "tractors")) %>%
  mutate(industry = case_match(ind_code, "110C" ~ "farms", "113F" ~ "forest, fish and serv")) %>%
  select(year, industry, asset, value) %>%
  arrange(industry, asset, year) %>%
  filter(year >= 1982) %>%
  mutate(value_norm = value / first(value), .by = c(industry, asset))

d %>%
  ggplot() +
  geom_line(aes(year, value)) + 
  scale_x_continuous(name = "", breaks = seq(1982, 2022, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale())) +
  facet_wrap(c("asset", "industry"), ncol = 2, scales = "free_y") +
  labs(title = "Asset stock")

d %>%
  ggplot() +
  geom_line(aes(year, value_norm)) +
  scale_x_continuous(name = "", breaks = seq(1982, 2022, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  facet_wrap(c("asset", "industry"), ncol = 2, scales = "fixed") +
  labs(title = "Asset stock, 1982 = 100%")

d %>%
  ggplot() +
  geom_line(aes(year, value_norm, color = asset, linetype = industry)) +
  scale_x_continuous(name = "", breaks = seq(1982, 2022, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale()))

# tractors and machine combined
d %>%
  summarize(value = sum(value), .by = c(year, industry)) %>%
  arrange(industry, year) %>%
  mutate(value_norm = value / first(value), .by = industry) %>%
  ggplot() +
  geom_line(aes(year, value_norm, linetype = industry)) +
  scale_x_continuous(name = "", breaks = seq(1982, 2022, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale()))

d %>%
  summarize(value = sum(value), .by = c(year, industry)) %>%
  arrange(industry, year) %>%
  mutate(value_norm = value / first(value), .by = industry) %>%
  filter(year == 2022)

```


# Machinery and labor shares


```{r}
#| fig-width: 4
#| fig-height: 3

d1 <- dat$farm_nat %>%
  filter(name %in% c("exp_labhire", "exp_labcont")) %>%
  bind_rows(dat$farm_labor_pre2002) %>%
  mutate(share = value / sum(value), .by = year) %>%
  filter(name == "exp_labcont") %>%
  select(year, share) %>%
  mutate(name = "paid labor")
d2 <- dat$fa %>%
  filter(year %in% 1982:2022) %>%
  summarize(value = sum(value), .by = c(year, ind_code)) %>%
  mutate(share = value / sum(value), .by = year) %>%
  filter(ind_code == "113F") %>%
  select(year, share) %>%
  mutate(name = "machinery")
paperdata$asset_sh <- d <- bind_rows(d1, d2)

d %>%
  filter(year %in% seq(1982, 2022, 5)) %>%
  pivot_wider(id_cols = year, values_from = share) %>%
  arrange(year)

d %>%
  ggplot() +
  geom_line(aes(year, share, linetype = name)) +
  scale_x_continuous(name = "", breaks = seq(1982, 2022, 10)) +
  scale_y_continuous(name = "", limits = c(0, 0.25), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  theme(legend.title = element_blank(), legend.position = "inside", legend.justification = c(0, 1), legend.position.inside = c(0, 1)) +
  labs(title = "Share managed by service providers")


```



# Crop heterogeneity

```{r}
df_crop <- agser$data_farm("county") %>%
  filter(year == 2017)
df_geo <- tigris::counties(cb = TRUE, resolution = "20m", year = 2017) %>%
  rename_with(tolower) %>%
  rename(stcty = geoid) %>%
  tigris::shift_geometry()
```

## map

```{r}
crops <- c("crop", "corn", "soy", "wheat", "veg", "fruit", "hort")
df <- df_crop %>%
  filter(name %in% paste0("sale_", crops)) %>%
  mutate(share = value / first(if_else(name == "sale_crop", value, NA), na_rm = TRUE), .by = stcty)

for (cr in crops[-1]) {
  p <- df_geo %>%
    left_join(df %>% filter(name == paste0("sale_", cr)) %>% select(!name), by = "stcty") %>%
    tm_shape() +
    tm_fill("share", col = NULL) +
    tm_title(cr)
  print(p)
}

```

## expenditure shares

```{r}
crops <- c("crop", "corn", "soy", "wheat", "veg", "fruit", "hort")
d1 <- df_crop %>%
  filter(name %in% paste0("sale_", crops)) %>%
  mutate(share = value / first(if_else(name == "sale_crop", value, NA), na_rm = TRUE), .by = stcty)

exps <- c("tot", "seed", "fert", "chem", "labhire", "labcont", "fuel", "repair", "agser_rent", "agser_cust", "agser_other")
d2 <- df_crop %>%
  filter(name %in% paste0("exp_", exps)) %>%
  mutate(share = value / first(if_else(name == "exp_tot", value, NA), na_rm = TRUE), .by = stcty)

d <- bind_rows(d1, d2) %>%
  filter(!name %in% c("sale_crop", "exp_tot")) %>%
  pivot_wider(id_cols = stcty, values_from = "share") %>%
  select(!stcty)

pairs(d, pch = ".")
cor(d, use = "pairwise.complete.obs")
```

## sales vs land

```{r}
crops <- c("crop", "corn", "soy", "wheat", "veg")
df <- df_crop %>%
  separate_wider_regex(name, c(var = ".+?", "_", com = ".+"), cols_remove = FALSE) %>%
  filter(var %in% c("land", "sale"), com %in% crops) %>%
  mutate(share = value / first(if_else(com == "crop", value, NA), na_rm = TRUE), .by = c(stcty, var))

df %>%
  pivot_wider(id_cols = c(stcty, com), names_from = var, values_from = value) %>%
  mutate(productivity = sale / land) %>%
  filter(is.finite(productivity)) %>%
  summarize(
    mean = mean(productivity),
    sd = sd(productivity),
    quantile(productivity) %>% as.list() %>% as_tibble(),
    .by = com
  )

df %>%
  filter(com != "crop") %>%
  pivot_wider(id_cols = c(stcty, com), names_from = var, values_from = share) %>%
  drop_na() %>%
  ggplot() +
  geom_point(aes(land, sale, color = com), size = 0.5)

df %>%
  pivot_wider(id_cols = c(stcty, com), names_from = var, values_from = value) %>%
  mutate(productivity = sale / land)
```


# Service providers

## est and pay trend

```{r}
paperdata$agser_nat <- d <- bind_rows(
  dat$rdc_2019 %>% mutate(pay = p941_sum + p943_sum, src = "LBD"),
  dat$agser_nat %>% mutate(src = "QCEW")
) %>%
  select(year, src, est, pay, pay_nag = p941_sum, pay_ag = p943_sum) %>%
  pivot_longer(!c(year, src)) %>%
  bind_rows(dat$rdc_2024$desc %>%
    filter(commodity == "crop", variable %in% c("est", "pay")) %>%
    mutate(value = nobs * mean) %>%
    select(year, name = variable, value) %>%
    mutate(value = if_else(name == "pay", value * 1000, value)) %>%
    mutate(src = "LBDr"))

d %>%
  ggplot() +
  geom_line(aes(year, value, color = src)) +
  scale_y_continuous(limits = c(0, NA)) +
  facet_wrap("name", scales = "free_y")

```



## QCEW suppression


```{r}
df_farm <- agser$data_farm("county") %>%
  filter(year == 2017, name %in% c("land_crop", "sale_crop", "n_sale_crop")) %>%
  pivot_wider(id_cols = stcty) %>%
  rename(farms = n_sale_crop, land = land_crop, sales = sale_crop)

df_qcew <- agser$qcew() %>%
  filter(year == 2017, agglvl_code %in% c("16", "76"), industry_code == "1151", own_code == "5") %>%
  select(stcty = area_fips, disc = disclosure_code, establishments = annual_avg_estabs, employment = annual_avg_emplvl, payroll = total_annual_wages) %>%
  collect() %>%
  mutate(disc = if_else(is.na(disc), "published", "suppressed"))


paperdata$qcew_sup <- d <- df_farm %>%
  left_join(df_qcew, by = "stcty") %>%
  mutate(disc = replace_na(disc, "no_agser")) %>%
  summarize(
    counties = n(),
    across(c(establishments, employment, payroll, farms, land, sales), \(x) sum(x, na.rm = TRUE)),
    .by = disc) %>%
  add_row(
    disc = "total",
    df_qcew %>% filter(stcty == "US000") %>% select(employment, payroll)
  ) %>%
  mutate(across(!disc, as.double)) %>%
  pivot_longer(!disc) %>%
  pivot_wider(id_cols = name, names_from = disc) %>%
  mutate(
    total = if_else(name %in% c("employment", "payroll"), total, no_agser + suppressed + published),
    suppressed = if_else(name %in% c("employment", "payroll"), total - published, suppressed),
    sh_suppressed = suppressed / total
  )

d
```






## size

-   Steady employment and payroll of non-agricultural worksers.
-   Doubling payroll of agricultural workers.
-   Consistent with greater increase of farm expense on contract labor and modest increase of other ag service expense.

```{r}
#| fig-width: 10
#| fig-height: 4
d <- dat$rdc_2019 %>%
  select(year, e941 = emp1_sum, p941 = p941_sum, p943_sum) %>%
  # filter(year >= 2002) %>%
  pivot_longer(!year)

d %>%
  ggplot() +
  geom_line(aes(year, value)) + 
  scale_x_continuous(name = "", breaks = seq(2002, 2017, 5)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  facet_wrap(c("name"), nrow = 1, scales = "free_y") +
  labs(title = "Employment and payroll (941 = non-ag, 943 = ag)")

d %>%
  arrange(year) %>%
  mutate(value = value / first(value), .by = c(name)) %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_x_continuous(name = "", breaks = seq(2002, 2017, 5)) +
  scale_y_continuous(name = "", limits = c(0.6, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale())) +
  facet_wrap(c("name"), ncol = 3, scales = "fixed") +
  labs(title = "Employment and payroll (941 = non-ag, 943 = ag), 2002 = 100%")
```


```{r}
paperdata$agser_size <- d <- dat$rdc_2019 %>%
  mutate(entry = birth / (birth + cont), exit = lag(death / (birth + cont))) %>%
  filter(year >= 2002) %>%
  select(year, est, entry, exit, pay_nonag = p941_sum, pay_ag = p943_sum) %>%
  pivot_longer(!year)

d %>%
  filter(name == "est") %>%
  ggplot() +
  geom_line(aes(year, value)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_number(scale_cut = scales::cut_short_scale()))

d %>%
  filter(name %in% c("entry", "exit")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_percent(scale_cut = scales::cut_short_scale()))


d %>%
  filter(name %in% c("pay_nonag", "pay_ag")) %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  scale_y_continuous(name = "", limits = c(0, NA), labels = scales::label_currency(scale_cut = scales::cut_short_scale()))


```

## county distribution

-   Most counties only have a few establishments and under \$1m of total payroll.
-   Market concentration has not changed from 2002 to 2017.

```{r}
#| fig-width: 10
#| fig-height: 3

paperdata$agser_distr <- d <- rdc_2024$desc %>%
  filter(commodity == "crop", variable %in% c("est", "pay", "pay_cr1")) %>%
  mutate(across(c(mean, sd, `25%`, `50%`, `75%`), \(x) if_else(variable == "pay", x * 1000, x)))

d %>%
  select(variable, year, nobs, mean, sd, `25%`, `50%`, `75%`) %>%
  arrange(variable, year)

yl <- list(est = c(0, 8), pay = c(0, 5e6), pay_cr1 = c(0,1))
ti <- list(est = "Establishments", pay = "Payroll", pay_cr1 = "Payroll share of largest est.")
ps <- list()
for (v in c("est", "pay", "pay_cr1")) {
  p <- d %>%
    filter(variable == v) %>%
    ggplot(aes(x = year)) +
    geom_boxplot(aes(lower = `25%`, upper = `75%`, middle = `50%`, ymin = `25%`, ymax = `75%`, group= year), stat = "identity", width = 2) +
    geom_line(aes(y = mean)) +
    geom_point(aes(y = mean)) +
    scale_x_continuous(breaks = PAR$lbd_years) +
    switch(
      v,
      est = scale_y_continuous(limits = yl[[v]], breaks = seq(0, 8, 2)),
      pay = scale_y_continuous(
        limits = yl[[v]], 
        labels = scales::label_currency(scale = 1000, scale_cut = scales::cut_short_scale())),
     pay_cr1 =  scale_y_continuous(limits = yl[[v]])
    ) +
    labs(title = ti[[v]], x = "", y = "")
  ps[[v]] <- p
}

ps[["est"]] + ps[["pay"]] + ps[["pay_cr1"]]
```


```{r}
d %>%
  filter(variable == "pay")  %>%
    ggplot(aes(x = year)) +
    geom_boxplot(aes(lower = `25%`, upper = `75%`, middle = `50%`, ymin = `25%`, ymax = `75%`, group= year), stat = "identity", width = 2) +
    geom_line(aes(y = mean)) +
    geom_point(aes(y = mean))
```


# Public data regressions


```{r}
df <- agser$data_farm("county")

```

## crop

```{r}
df_reg <- df %>%
  filter(name %in% c("land_crop", "sale_crop", "n_sale_crop", "exp_tot", "exp_labhire", "exp_labcont", "exp_agser_cust", "asset_mach")) %>%
  pivot_wider(id_cols = c(year, stcty)) %>%
  mutate(
    yield_crop = if_else(land_crop > 0 & sale_crop > 0, sale_crop / land_crop, NA),
    sh_labcont = exp_labcont / (exp_labhire + exp_labcont)
  ) %>%
  mutate(across(c("yield_crop", "land_crop", "exp_tot", "exp_labhire", "exp_labcont", "exp_agser_cust", "asset_mach"), \(x) log(x + 1)))


result_reg <-
  # create a table of estimation specification combinations
  expand_grid(
    year = seq(2002, 2022, 5)
  ) %>%
  # iterate over spec table rows, construct model formula and estimate
  pmap(function(...) {
    row <- list(...)
    d <- df_reg %>%
      filter(year == row$year)
    f <- "yield_crop ~ land_crop + exp_tot + sh_labcont + asset_mach + exp_agser_cust"
    # f <- "yield_crop ~ land_crop + exp_labhire + exp_labcont + exp_agser_cust"
    m <- lm_robust(as.formula(f), data = d, weights = d$n_sale_crop)
    # return single-row table with specs, formula and model object
    tibble(..., formula = f, model = list(m))
  }) %>%
  bind_rows()


result_reg$model %>%
  set_names(result_reg$year) %>%
  modelsummary(stars = TRUE, output = "huxtable")

```

## corn

```{r}
df_reg <- df %>%
  filter(name %in% c("land_corn", "land_crop", "prod_corn", "n_land_corn", "exp_tot", "exp_labhire", "exp_labcont", "exp_agser_cust", "asset_mach")) %>%
  pivot_wider(id_cols = c(year, stcty)) %>%
  mutate(
    yield_corn = if_else(land_corn > 0 & prod_corn > 0, prod_corn / land_corn, NA),
    sh_labcont = exp_labcont / (exp_labhire + exp_labcont),
    sh_land_corn = land_corn / land_crop
  ) %>%
  mutate(across(c("yield_corn", "land_corn", "exp_tot", "exp_labhire", "exp_labcont", "exp_agser_cust", "asset_mach", "prod_corn"), \(x) log(x + 1)))


result_reg <-
  # create a table of estimation specification combinations
  expand_grid(
    year = seq(2002, 2022, 5)
  ) %>%
  # iterate over spec table rows, construct model formula and estimate
  pmap(function(...) {
    row <- list(...)
    d <- df_reg %>%
      filter(year == row$year)
    # f <- "land_corn ~ exp_labhire + exp_labcont + asset_mach + exp_agser_cust"
    # f <- "yield_corn ~ land_corn + exp_labhire + exp_labcont + asset_mach + exp_agser_cust"
    # f <- "yield_corn ~ land_corn + exp_tot + sh_labcont + asset_mach + exp_agser_cust"
    # m <- lm_robust(as.formula(f), data = d, weights = d$n_land_corn)
    m <- lm_robust(as.formula(f), data = d, weights = d$sh_land_corn)
    # m <- lm_robust(as.formula(f), data = d)
    # return single-row table with specs, formula and model object
    tibble(..., formula = f, model = list(m))
  }) %>%
  bind_rows()


result_reg$model %>%
  set_names(result_reg$year) %>%
  modelsummary(stars = TRUE, output = "huxtable")

# state fixed effects
# fixed panel of countieы
# total output as outcome

```



## soy

```{r}
df_reg <- df %>%
  filter(name %in% c("land_soy", "prod_soy", "n_land_soy", "exp_tot", "exp_labhire", "exp_labcont", "exp_agser_cust", "asset_mach")) %>%
  pivot_wider(id_cols = c(year, stcty)) %>%
  mutate(
    yield_soy = if_else(land_soy > 0 & prod_soy > 0, prod_soy / land_soy, NA),
    sh_labcont = exp_labcont / (exp_labhire + exp_labcont)
  ) %>%
  mutate(across(c("yield_soy", "land_soy", "exp_tot", "exp_labhire", "exp_labcont", "exp_agser_cust", "asset_mach"), \(x) log(x + 1)))


result_reg <-
  # create a table of estimation specification combinations
  expand_grid(
    year = seq(2002, 2022, 5)
  ) %>%
  # iterate over spec table rows, construct model formula and estimate
  pmap(function(...) {
    row <- list(...)
    d <- df_reg %>%
      filter(year == row$year)
    f <- "yield_soy ~ land_soy + exp_labhire + exp_labcont + asset_mach + exp_agser_cust"
    # f <- "yield_soy ~ land_soy + exp_tot + sh_labcont + asset_mach + exp_agser_cust"
    m <- lm_robust(as.formula(f), data = d, weights = d$n_land_soy)
    # return single-row table with specs, formula and model object
    tibble(..., formula = f, model = list(m))
  }) %>%
  bind_rows()


result_reg$model %>%
  set_names(result_reg$year) %>%
  modelsummary(stars = TRUE, output = "huxtable")


```



## wheat

```{r}
df_reg <- df %>%
  filter(name %in% c("land_wheat", "prod_wheat", "n_land_wheat", "exp_tot", "exp_labhire", "exp_labcont", "exp_agser_cust", "asset_mach")) %>%
  pivot_wider(id_cols = c(year, stcty)) %>%
  mutate(
    yield_wheat = if_else(land_wheat > 0 & prod_wheat > 0, prod_wheat / land_wheat, NA),
    sh_labcont = exp_labcont / (exp_labhire + exp_labcont)
  ) %>%
  mutate(across(c("yield_wheat", "land_wheat", "exp_tot", "exp_labhire", "exp_labcont", "exp_agser_cust", "asset_mach"), \(x) log(x + 1)))


result_reg <-
  # create a table of estimation specification combinations
  expand_grid(
    year = seq(2002, 2022, 5)
  ) %>%
  # iterate over spec table rows, construct model formula and estimate
  pmap(function(...) {
    row <- list(...)
    d <- df_reg %>%
      filter(year == row$year)
    f <- "yield_wheat ~ land_wheat + exp_labhire + exp_labcont + asset_mach + exp_agser_cust"
    # f <- "yield_wheat ~ land_wheat + exp_tot + sh_labcont + asset_mach + exp_agser_cust"
    m <- lm_robust(as.formula(f), data = d, weights = d$n_land_wheat)
    # return single-row table with specs, formula and model object
    tibble(..., formula = f, model = list(m))
  }) %>%
  bind_rows()


result_reg$model %>%
  set_names(result_reg$year) %>%
  modelsummary(stars = TRUE, output = "huxtable")


```

# LBD regressions

Do counties with larger presence of service providers (measured by payroll) have higher agricultural productivity?

"Crop yield" is sales per acre.
Corn, wheat and soy are production per acre.

## static productivity

Regression model:

```         
logyield ~ logpay * (logland + logavgland) + hhi_land
```

### all crops, sales per acre, 2017

-   AgSer payroll elasticity = 0.2
-   Controlling for farmland size and distribution
-   Relationship is stronger in counties with more farmland and smaller farms

```{r}
paperdata$reg_static <- rdc_2024$static_yield

models <- rdc_2024$static_yield %>%
  filter(commodity == "crop", year == 2017)

fits <- unique(models$model) %>%
  map(\(l) filter(models, model == l))

modelsummary(fits, output = "huxtable")

```

### logpay coef by commodity-year

-   Strongest effect for total crop productivity
-   Weaker, but positive for corn and wheat
-   Weakest for soybeans

```{r}
models <- rdc_2024$static_yield %>%
  filter(model == "logyield ~ logpay + logland + logavgland + hhi_land",
         term == "logpay") %>%
  rename(label = sample_label)

models %>%
  left_join(models %>% tidy() %>% select(label, conf.low, conf.high),
            by = "label") %>%
  ggplot(aes(year)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = commodity), alpha = 0.2) +
  geom_line(aes(y = estimate, color = commodity)) +
  scale_x_continuous(breaks = PAR$lbd_years)


```





### interacted

Land interaction: Positive effect on productivity is amplified in larger counties for all except corn.

Average land interaction: For crops, positive effect on productivity is amplified in counties with smaller average land (more farms).


```{r}

paperdata$reg_int <- models <- expand_grid(commodity = c("corn", "soy", "wheat"), year = seq(2002, 2017, 5)) %>%
  pmap(function(commodity, year) {
    model <- list()
    model[[1]] <- rdc_2024$static_yield %>%
      filter(
        model == "logyield ~ logpay * (logland + logavgland) + hhi_land",
        commodity == !!commodity, 
        year == !!year)
    tibble(commodity, year, label = paste0(commodity, "_", year), model)
  }) %>%
  bind_rows()

for (x in c("corn", "soy", "wheat")) {
  models %>%
    filter(commodity == x) %>%
    pull(model, name = label) %>%
    modelsummary(
      output = "huxtable",
      coef_map = c("logpay" = "Payroll", "logpay:logland" = "Payroll*Land", "logpay:logavgland" = "Payroll*AvgLand", "logland" = "Land", "logavgland" = "AvgLand", "hhi_land" = "LandHHI")
    ) %>%
    print()
}


```

#### marginal effects

```{r}
# model equation:
# logyield ~ b0 + b1 logpay + b2 logpay*logland + b3 logland + b4 logpay*logavgland + b5 logavgland
# marinal effect:
# d(logyield) / d(logpay) = b1 + b2 * logland + b4 * logavgland

d_coef <- rdc_2024$static_yield %>% 
  filter(model == "logyield ~ logpay * (logland + logavgland) + hhi_land", str_detect(term, "logpay")) %>%
  pivot_wider(id_cols = c(year, commodity), names_from = term, values_from = estimate)

d_land <- agser$data_farm("county") %>%
  separate_wider_regex(name, c(measure = ".+", "_", commodity = ".+")) %>%
  filter(year <= 2017, measure %in% c("land", "n_land"), commodity %in% c("crop", "corn", "soy", "wheat")) %>%
  pivot_wider(id_cols = c(year, commodity, stcty), names_from = measure) %>%
  mutate(avgland = land / n_land) %>%
  mutate(logland = log(land), logavgland = log(avgland))

d_me <- inner_join(d_coef, d_land, by = c("year", "commodity")) %>%
  mutate(me = logpay + `logpay:logland` * logland + `logpay:logavgland` * logavgland)
d_me
```


```{r}
f <- list(yr = 2012, com = "wheat")

d_coef %>%
  filter(year == f$yr, commodity == f$com)

d_me %>%
  filter(year == f$yr, commodity == f$com) %>%
  ggplot() +
  geom_point(aes(logland, me))

d_me %>%
  filter(year == f$yr, commodity == f$com) %>%
  ggplot() +
  geom_point(aes(logavgland, me))
  
d_me %>%
  filter(year == f$yr, commodity == f$com) %>%
  ggplot() +
  geom_point(aes(logland, logavgland, color = me)) +
  scale_color_gradient(low = "blue", high = "green")


d_me %>%
  filter(commodity != "crop") %>%
  ggplot() +
  geom_point(aes(logland, logavgland, color = me)) +
  scale_color_gradient(low = "#5e3c99", high = "#e66101") +
  facet_wrap(c("commodity", "year"))

```


```{r}
#' var-cov matrix from static yield model with interactions
rdc_vcov <- function(commodity, year) {
  model <- "logyield ~ logpay * (logland + logavgland) + hhi_land"
  coefs <- c("logpay", "logpay:logland", "logpay:logavgland")
  vc <- matrix(nrow = 3, ncol = 3, dimnames = list(coefs, coefs))
  se <- rdc_2024$static_yield %>%
    filter(commodity == !!commodity, year == !!year, model == !!model, term %in% coefs) %>%
    pull(std.error, name = term)
  se <- se[coefs]
  diag(vc) <- se^2
  df_cov <- rdc_2024$static_yield %>%
    filter(commodity == !!commodity, year == !!year, model == !!model) %>%
    filter(term == "coef_covar")
  for (i in 1:nrow(df_cov)) {
    t1 <- df_cov$term_1[i]
    t2 <- df_cov$term_2[i]
    vc[t1, t2] <- vc[t2, t1] <- df_cov$covariance[i]
  }
  vc
}


df <- rdc_2024$static_yield %>%
  filter(model == "logyield ~ logpay * (logland + logavgland) + hhi_land",
         str_detect(term, "logpay")) %>%
  select(commodity, year, sample_label, term, estimate) %>%
  pivot_wider(id_cols = c(commodity, year, sample_label), names_from = term, values_from = estimate) %>%
  rename_with(\(x) str_replace(x, ":", "_")) %>%
  left_join(
    rdc_2024$desc %>%
      select(sample_label, variable, "25%", "50%", "75%") %>%
      filter(variable %in% c("land", "avgland")) %>%
      mutate(variable = paste0("log", variable),
             across(c("25%", "50%", "75%"), log)) %>%
      pivot_wider(id_cols = sample_label, names_from = variable, values_from = c("25%", "50%", "75%"), names_glue = "{variable}_{.value}") %>%
      rename_with(\(x) str_sub(x, 1, -2), .cols = ends_with("%")),
    by = "sample_label") %>%
  pmap(function(...) {
    row <- list(...)
    vc <- rdc_vcov(row$commodity, row$year)
    tibble(..., vc = list(vc))
  }) %>%
  bind_rows() %>%
  rowwise() %>%
  mutate(
    dydx_logland_25 = logpay + logpay_logland * logland_25 + logpay_logavgland * logavgland_50,
    dydx_logland_75 = logpay + logpay_logland * logland_75 + logpay_logavgland * logavgland_50,
    dydx_logavgland_25 = logpay + logpay_logland * logland_50 + logpay_logavgland * logavgland_25,
    dydx_logavgland_75 = logpay + logpay_logland * logland_50 + logpay_logavgland * logavgland_75
  )

df
```



```{r}
d <- df %>%
  select(commodity, year, starts_with("dydx_")) %>%
  pivot_longer(!c(commodity, year), names_pattern = "dydx_(.+)_(..)", names_to = c("atvar", "atval"), values_to = "dydx") %>%
  mutate(atval = paste0(atval, "%"))
  
d %>%
  filter(atvar == "logland") %>%
  rename(land = atval) %>%
  ggplot() +
  geom_line(aes(year, dydx, color = land)) +
  facet_wrap("commodity", scale = "free_y") +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  scale_x_continuous(breaks = PAR$lbd_years) +
  ylab("marginal effect of logpay")


d %>%
  filter(atvar == "logavgland") %>%
  rename(avg_land = atval) %>%
  ggplot() +
  geom_line(aes(year, dydx, color = avg_land)) +
  facet_wrap("commodity", scale = "free_y") +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  scale_x_continuous(breaks = PAR$lbd_years) +
  ylab("marginal effect of logpay")


```




```{r}

#' Std error of marginal effect of logpay at given values of logland and logavgland
me_se <- function(vc, land, avgland) {
  jac <- matrix(c(1, land, avgland), ncol = 1)
  (t(jac) %*% vc %*% jac) %>%
    c() %>%
    sqrt()
}

# d_land %>%
#   filter(commodity != "crop", year == 2012) %>%
#   group_by(commodity) %>%
#   skim(logland, logavgland)

df_land_p50 <- rdc_2024$desc %>%
  filter(variable %in% c("land", "avgland")) %>%
  mutate(value = log(`50%`)) %>%
  pivot_wider(id_cols = c(commodity, year), names_from = variable, names_glue = "log{variable}_p50")

paperdata$reg_me <- d <- df %>%
  filter(commodity != "crop") %>%
  left_join(df_land_p50, by = c("commodity", "year")) %>%
  pmap(function(commodity, year, logpay, logpay_logland, logpay_logavgland, vc, logland_p50, logavgland_p50, ...) {
    
    bind_rows(
      tibble(atvar = "land", logland = seq(2, 11), logavgland = logavgland_p50),
      tibble(atvar = "avgland", logland = logland_p50, logavgland = seq(2, 6))
    ) %>%
    rowwise() %>%
    mutate(
      commodity = commodity,
      year = year,
      dydx = logpay + logpay_logland * logland + logpay_logavgland * logavgland,
      se = me_se(vc, logland, logavgland),
      lo = dydx - 1.96 * se,
      hi = dydx + 1.96 * se
    )
  }) %>%
  bind_rows()

d %>%
  filter(atvar == "land") %>%
  ggplot() +
  geom_ribbon(aes(logland, ymin = lo, ymax = hi), alpha = 0.2) +
  geom_line(aes(logland, dydx)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  facet_wrap(c("commodity", "year"), scales = "free_y")

d %>%
  filter(atvar == "avgland") %>%
  ggplot() +
  geom_ribbon(aes(logavgland, ymin = lo, ymax = hi), alpha = 0.2) +
  geom_line(aes(logavgland, dydx)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  facet_wrap(c("commodity", "year"), scales = "free_y")

```



## productivity growth

Regression model (`diff_logyield = logyield(y+5) - logyield(y)`):

```         
diff_logyield ~ logpay + logyield + logland + logavgland + hhi_land
```

-   Counties with more AgSer payroll tend to have higher increases in productivity over the following 5 years.
-   Strongest effect in 2007.

```{r}

paperdata$reg_dynamic <- rdc_2024$dynamic_yield

models <- rdc_2024$dynamic_yield %>%
  mutate(label = paste0(commodity, "_", year), .before = 1)

fits <- unique(models$label) %>%
  set_names() %>%
  map(\(l) filter(models, label == l))

for (x in PAR$commodities) {
  models %>%
    filter(commodity == x) %>%
    distinct(label) %>%
    pull() %>%
    `[`(fits, .) %>%
    modelsummary(output = "huxtable") %>%
    print()
}

```


